---
title: "Project 3"
author: "Lucas Andersen, Arjun Gautam, Cameron Her, Efrain Olivares"
date: "2025-08-03"
output: pdf_document
---

# PREPARING THE DATA

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)

haititrain <- read.csv("https://gedeck.github.io/DS-6030/project/HaitiPixels.csv")

```

```{r}
setwd("/Users/arjungautam/Documents/UVA/Summer25/DS6030/HoldOutData")

process_load_holdout_file <- function(filename) {
    fileLines <- readLines(sprintf(filename), skip=7)
    fileLines[8] <- gsub(";", "", fileLines[8])
    fileLines[8] <- gsub("Map X", "Map_X", fileLines[8])
    fileLines[8] <- gsub("Map Y", "Map_Y", fileLines[8])
    writeLines(fileLines, sprintf(filename))
    read_table(sprintf(filename), skip=7)
}

rename_set_target <- function(df, class_target) {
    colnames(df)[colnames(df) == 'B1'] <- 'Red'
    colnames(df)[colnames(df) == 'B2'] <- 'Green'
    colnames(df)[colnames(df) == 'B3'] <- 'Blue'
    df %>%mutate(target = class_target)
}

# collect holdout blue tarp data
holdout_blue_tarp <- bind_rows(
    process_load_holdout_file("orthovnir067_ROI_Blue_Tarps.txt"),
    process_load_holdout_file("orthovnir069_ROI_Blue_Tarps.txt"),
    process_load_holdout_file("orthovnir078_ROI_Blue_Tarps.txt")
)

test_blue_tarp <- holdout_blue_tarp %>%
    rename_set_target('blue_tarp') %>%
    dplyr::select(c(Red, Green, Blue, target))

head(test_blue_tarp)
nrow(test_blue_tarp)

```


```{r}
# collect holdout non_blue tarp data
holdout_non_blue_tarp <- bind_rows(
  process_load_holdout_file("/Users/arjungautam/Documents/UVA/Summer25/DS6030/HoldOutData/orthovnir078_ROI_NON_Blue_Tarps.txt"),
  process_load_holdout_file("/Users/arjungautam/Documents/UVA/Summer25/DS6030/HoldOutData/orthovnir067_ROI_NOT_Blue_Tarps.txt"),
  process_load_holdout_file("/Users/arjungautam/Documents/UVA/Summer25/DS6030/HoldOutData/orthovnir069_ROI_NOT_Blue_Tarps.txt"),
  process_load_holdout_file("/Users/arjungautam/Documents/UVA/Summer25/DS6030/HoldOutData/orthovnir057_ROI_NON_Blue_Tarps.txt")
)

test_non_blue_tarp <- holdout_non_blue_tarp %>%
    rename_set_target('non_blue') %>%
    dplyr::select(c(Red, Green, Blue, target))

#head(holdout_non_blue_tarp)
nrow(test_non_blue_tarp)
```

```{r}
# combine both for final set
holdout_class <- bind_rows(
  test_blue_tarp,
  test_non_blue_tarp
)
```

```{r}
haititrain <- haititrain %>%
  mutate(
    Class = if_else(Class == "Blue Tarp", "Blue Tarp", "Non-Tarp"),
    Class = factor(Class)
  )
```

```{r}
holdout_class <- holdout_class %>% rename(Class = target)
```

```{r}

new_datapoints <- read_table('image_rgb.txt') 
colnames(new_datapoints) <- c("Red", "Green", "Blue") 
thepreds <- predict(lda_fit, new_data=new_datapoints) 
write.csv(thepreds, "LDAthepreds.csv", row.names = FALSE) 
```

```{r}
new_datapoints <- read_table('image_rgb.txt') 
colnames(new_datapoints) <- c("Red", "Green", "Blue") 
thepreds <- predict(log_fit, new_data=new_datapoints) 
write.csv(thepreds, "LOGthepreds.csv", row.names = FALSE) 
```

#LDA, QDA, LOG REG


```{r}
pixel_recipe <- recipe(Class ~ Red + Green + Blue, data = haititrain) %>%
  step_normalize(all_predictors())
```

```{r}
library(discrim)

# Logistic Regression
log_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# LDA
lda_spec <- discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

# QDA
qda_spec <- discrim_quad() %>%
  set_engine("MASS") %>%
  set_mode("classification")

log_wf <- workflow() %>% add_model(log_spec) %>% add_recipe(pixel_recipe)
lda_wf <- workflow() %>% add_model(lda_spec) %>% add_recipe(pixel_recipe)
qda_wf <- workflow() %>% add_model(qda_spec) %>% add_recipe(pixel_recipe)

log_fit <- fit(log_wf, data = haititrain)
lda_fit <- fit(lda_wf, data = haititrain)
qda_fit <- fit(qda_wf, data = haititrain)


```

```{r}
library(tidymodels)

set.seed(123)
resamples <- vfold_cv(haititrain, v = 10, strata = Class)
custom_metrics <- metric_set(accuracy, roc_auc)

pixel_recipe <- recipe(Class ~ Red + Green + Blue, data = haititrain) %>% step_normalize(all_predictors())

log_spec <- logistic_reg() %>% set_engine("glm") %>% set_mode("classification")


log_wf <- workflow() %>% add_model(log_spec) %>% add_recipe(pixel_recipe)

cv_control <- control_resamples(save_pred = TRUE)

log_cv <- fit_resamples(
  log_wf,
  resamples = resamples,
  metrics = custom_metrics,
  control = cv_control
)

lda_cv <- fit_resamples(
  lda_wf,
  resamples = resamples,
  metrics = custom_metrics,
  control = cv_control
)

qda_cv <- fit_resamples(
  qda_wf,
  resamples = resamples,
  metrics = custom_metrics,
  control = cv_control
)


cv_metrics <- bind_rows(
collect_metrics(log_cv) %>% mutate(model="Log"),
collect_metrics(lda_cv) %>% mutate(model="LDA"),
collect_metrics(qda_cv) %>% mutate(model="QDA")
)

cv_metrics %>%
dplyr::select(model, .metric, mean) %>%
pivot_wider(names_from=.metric, values_from=mean) %>%
knitr::kable(caption="Cross-validation performance metrics", digits=5)

```
```{r}
roc_cv_plot <- function(model_cv, model_name) {
cv_predictions <- collect_predictions(model_cv)
cv_roc <- cv_predictions %>%
roc_curve(truth=Class, .pred_Class, event_level="first")
g <- autoplot(cv_roc) +
labs(title=model_name)
return(g)
}

bind_rows(
collect_predictions(log_cv) %>% mutate(model="Logistic regression"),
collect_predictions(lda_cv) %>% mutate(model="LDA"),
collect_predictions(qda_cv) %>% mutate(model="QDA")
) %>%
group_by(model) %>%
  roc_curve(truth = Class, `.pred_Blue Tarp`, event_level = "first") %>%
  autoplot() +
  labs(title = "ROC Curves on Training Data", color = "Model")
```



```{r}
library(yardstick)
library(probably)

get_j_curve <- function(fitted_model, data, model_name) {
  probs <- predict(fitted_model, new_data = haititrain, type = "prob")
  truth_df <- data.frame(Class = haititrain$Class)  

  combined <- bind_cols(probs, truth_df)

  threshold_perf(
    combined,
    truth = Class,
    estimate = `.pred_Blue Tarp`,
    thresholds = seq(0.05, 0.95, 0.01),
    metrics = metric_set(j_index),
    event_level = "first"
  ) %>%
    mutate(model = model_name)
}

j_logreg <- get_j_curve(log_fit, haititrain, "Logistic")
j_lda    <- get_j_curve(lda_fit, haititrain, "LDA")
j_qda    <- get_j_curve(qda_fit, haititrain, "QDA")

j_all <- bind_rows(j_logreg, j_lda, j_qda)

ggplot(j_all, aes(x = .threshold, y = .estimate, color = model)) +
  geom_line(size = 1) +
  geom_vline(
    data = j_all %>% group_by(model) %>% slice_max(.estimate),
    aes(xintercept = .threshold),
    linetype = "dashed"
  ) +
  labs(
    title = "J-index vs. Threshold (Haiti Blue Tarp Detection)",
    x = "Threshold",
    y = "J Index"
  ) +
  theme_minimal()
```

```{r}
performance_all <- function(model, data, model_name) {

  probs <- predict(model, data, type = "prob") %>% bind_cols(data.frame(Class = haititrain$Class)) 

  threshold_perf(
  probs,
  truth = Class,
  estimate = `.pred_Blue Tarp`,
  thresholds = seq(0.05, 0.95, 0.01),
  metrics = metric_set(j_index, accuracy, sens, spec,f_meas),
  event_level = "first"
) %>%
    mutate(model = model_name)
}

performance_logreg <- performance_all(log_fit, haititrain, "Logistic-Train")
performance_lda    <- performance_all(lda_fit, haititrain, "LDA-Train")
performance_qda    <- performance_all(qda_fit, haititrain, "QDA-Train")

performance_all_threshold <- bind_rows(performance_logreg, performance_lda, performance_qda)

best_thresholds <- performance_all_threshold %>%
  filter(.metric == "j_index") %>%
  group_by(model) %>%
  slice_max(.estimate, with_ties = FALSE) %>%
  dplyr::select(model, best_threshold = .threshold)


final_metrics <- performance_all_threshold %>%
  inner_join(best_thresholds, by = "model") %>%
  filter(.threshold == best_threshold) %>%
  dplyr::select(model, metric = .metric, threshold = .threshold, estimate = .estimate) %>%
  pivot_wider(names_from = metric, values_from = estimate)

final_metrics

```

```{r}
final_metrics %>%
  arrange(desc(j_index)) %>%
  kable(
    caption = "Final Performance Metrics at Optimal Threshold (J-index)",
    digits = 3,
    col.names = c("Model", "Threshold", "Accuracy", "J-index", "Sensitivity", "Specificity", "F_meas")
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```


```{r}
lda_thresh <- best_thresholds %>% filter(model == "LDA-Train") %>% pull(best_threshold)
log_thresh <- best_thresholds %>% filter(model == "Logistic-Train") %>% pull(best_threshold)
qda_thresh <- best_thresholds %>% filter(model == "QDA-Train") %>% pull(best_threshold)

predict_holdout_thresholded <- function(model_fit, threshold, model_name) {
  predict(model_fit, new_data = holdout_class, type = "prob") %>%
    rename(.pred_Blue_Tarp = `.pred_Blue Tarp`) %>%
    bind_cols(holdout_class) %>%
    mutate(
      pred_class = if_else(.pred_Blue_Tarp > threshold, "blue_tarp", "non_blue"),
      model = model_name
    )
}

log_preds <- predict_holdout_thresholded(log_fit, log_thresh, "Logistic")
lda_preds <- predict_holdout_thresholded(lda_fit, lda_thresh, "LDA")
qda_preds <- predict_holdout_thresholded(qda_fit, qda_thresh, "QDA")

log_thresh <- best_thresholds %>% filter(model == "Logistic-Train") %>% pull(best_threshold)
lda_thresh <- best_thresholds %>% filter(model == "LDA-Train") %>% pull(best_threshold)
qda_thresh <- best_thresholds %>% filter(model == "QDA-Train") %>% pull(best_threshold)

log_preds <- predict_holdout_thresholded(log_fit, log_thresh, "Logistic")
lda_preds <- predict_holdout_thresholded(lda_fit, lda_thresh, "LDA")
qda_preds <- predict_holdout_thresholded(qda_fit, qda_thresh, "QDA")

holdout_preds_all<-bind_rows(log_preds, lda_preds, qda_preds)

holdout_metrics <- holdout_preds_all %>%
  mutate(
    pred_class = recode(pred_class,
                        "blue_tarp" = "Blue Tarp",
                        "non_blue" = "Non-Tarp"),
    pred_class = factor(pred_class, levels = c("Non-Tarp", "Blue Tarp")),
    Class = factor(Class, levels = c("Non-Tarp", "Blue Tarp"))
  ) %>%
  group_by(model) %>%
  summarise(
    accuracy    = accuracy_vec(truth = Class, estimate = pred_class),
    sensitivity = sens_vec(truth = Class, estimate = pred_class),
    specificity = spec_vec(truth = Class, estimate = pred_class),
    f_meas      = f_meas_vec(truth = Class, estimate = pred_class, event_level = "first"),
    j_index     = sensitivity + specificity - 1,
    .groups     = "drop"
  )

holdout_metrics
```

```{r}
library(dplyr)
library(ggplot2)

lda_preds_holdout <- holdout_class %>%
  bind_cols(predict(lda_fit, new_data = holdout_class))

lda_preds_holdout <- lda_preds_holdout %>%
  mutate(
    color = rgb(Red, Green, Blue, maxColorValue = 255)
  )

predicted_blue_plot <- lda_preds_holdout %>%
  filter(.pred_class == "Blue Tarp")

g3 <- ggplot(predicted_blue_plot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "LDA - Predicted Blue Tarp Colors (Test Set)")

incorrect_blue_plot <- lda_preds_holdout %>%
  filter(.pred_class == "Blue Tarp" & Class == "Non-Tarp")

g4 <- ggplot(incorrect_blue_plot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "LDA - Incorrectly Predicted Blue Tarp Colors (Test Set)")


g3
g4

```

```{r}

qda_preds_holdout <- holdout_class %>%
  bind_cols(predict(qda_fit, new_data = holdout_class)) %>%
  mutate(color = rgb(Red, Green, Blue, maxColorValue = 255))


qda_predicted_blue <- qda_preds_holdout %>%
  filter(.pred_class == "Blue Tarp")

g3_qda <- ggplot(qda_predicted_blue, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "QDA - Predicted Blue Tarp Colors (Test Set)")


qda_incorrect_blue <- qda_preds_holdout %>%
  filter(.pred_class == "Blue Tarp" & Class == "Non-Tarp")

g4_qda <- ggplot(qda_incorrect_blue, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "QDA - Incorrectly Predicted Blue Tarp Colors (Test Set)")


log_preds_holdout <- holdout_class %>%
  bind_cols(predict(log_fit, new_data = holdout_class)) %>%
  mutate(color = rgb(Red, Green, Blue, maxColorValue = 255))


log_predicted_blue <- log_preds_holdout %>%
  filter(.pred_class == "Blue Tarp")

g3_log <- ggplot(log_predicted_blue, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "Logistic Regression - Predicted Blue Tarp Colors (Test Set)")


log_incorrect_blue <- log_preds_holdout %>%
  filter(.pred_class == "Blue Tarp" & Class == "Non-Tarp")

g4_log <- ggplot(log_incorrect_blue, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal() +
  labs(title = "Logistic Regression - Incorrectly Predicted Blue Tarp Colors (Test Set)")

g3_qda
g4_qda
g3_log
g4_log
```



```{r}
holdout_metrics %>%
  arrange(desc(j_index)) %>%
  kable(
    caption = "Performance Metrics at Optimal Threshold on Test Data",
    digits = 5,
    col.names = c("Model", "Accuracy", "Sensitivity", "Specificity","F_meas", "J-index")
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
holdout_probs_all <- bind_rows(log_preds, lda_preds, qda_preds)


holdout_probs_all <- holdout_probs_all %>%
  mutate(Class = factor(Class, levels = c("Non-Tarp", "Blue Tarp")))


roc_data <- holdout_probs_all %>%
  group_by(model) %>%
  roc_curve(truth = Class, .pred_Blue_Tarp, event_level = "second")

autoplot(roc_data) +
  labs(
    title = "ROC Curves on Test Data",
    x = "1 - Specificity",
    y = "Sensitivity",
    color = "Model"
  ) +
  theme_minimal()
```

```{r}
roc_auc_results <- holdout_probs_all %>%
  group_by(model) %>%
  roc_auc(truth = Class, .pred_Blue_Tarp, event_level = "second") %>%
  ungroup() %>%
  select(model, roc_auc = .estimate)

roc_auc_results %>%
  mutate(roc_auc = round(roc_auc, 5)) %>%
  kable(
    caption = "ROC AUC Values for Each Model on Test Set",
    col.names = c("Model", "ROC AUC"),
    digits = 5
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

```{r}
pca_recipe_train <- recipe(Class ~ Red + Green + Blue, data = haititrain) %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 2) %>% prep()

pca_pred_train <- bake(pca_recipe_train, new_data = haititrain)


logreg_2d <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  fit(Class ~ PC1 + PC2, data = pca_pred_train)

grid <- expand.grid(
  PC1 = seq(min(pca_pred_train$PC1), max(pca_pred_train$PC1), length.out = 100),
  PC2 = seq(min(pca_pred_train$PC2), max(pca_pred_train$PC2), length.out = 100)
)

grid_preds <- grid %>%
  mutate(pred = predict(logreg_2d, new_data = grid)$.pred_class)

ggplot(pca_pred_train, aes(x = PC1, y = PC2, color = Class)) +
  geom_point(alpha = 0.6) +
  geom_contour(data = grid_preds, aes(z = as.numeric(pred == "Blue Tarp")),
               breaks = 0.5, color = "black") +
  labs(title = "Logistic Regression Decision Boundary on Training Data") +
  theme_minimal()
```
```{r}
lda_fit_2d <- discrim_linear() %>%
  set_engine("MASS") %>%
  fit(Class ~ PC1 + PC2, data = pca_pred_train)

grid_vals <- expand.grid(
  PC1 = seq(min(pca_pred_train$PC1), max(pca_pred_train$PC1), length.out = 200),
  PC2 = seq(min(pca_pred_train$PC2), max(pca_pred_train$PC2), length.out = 200)
)

grid_vals$Class <- predict(lda_fit_2d, new_data = grid_vals)$.pred_class


ggplot(pca_pred_train, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = Class), alpha = 0.4) +
  geom_contour(data = grid_vals, aes(z = as.numeric(Class == "Blue Tarp")), 
               breaks = 0.5, color = "black") +
  labs(title = "LDA Decision Boundary on Training Data", color = "Class") +
  theme_minimal()
```
```{r}
qda_fit_2d <- discrim_quad() %>%
  set_engine("MASS") %>%
  fit(Class ~ PC1 + PC2, data = pca_pred_train)

grid_vals_qda <- grid_vals %>%
  select(PC1, PC2) %>%
  mutate(Class = predict(qda_fit_2d, new_data = .)$.pred_class)

ggplot(pca_pred_train, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = Class), alpha = 0.4) +
  geom_contour(data = grid_vals_qda, aes(z = as.numeric(Class == "Blue Tarp")), 
               breaks = 0.5, color = "black") +
  labs(title = "QDA Decision Boundary on Training Data", color = "Class") +
  theme_minimal()

```
```{r}
grid_vals <- expand.grid(
  Red   = seq(min(holdout_class$Red), max(holdout_class$Red), length.out = 100),
  Green = seq(min(holdout_class$Green), max(holdout_class$Green), length.out = 100),
  Blue  = seq(min(holdout_class$Blue), max(holdout_class$Blue), length.out = 100)
)

grid_vals <- grid_vals %>%
  mutate(across(everything(), as.integer))

grid_preds_lda <- grid_vals %>%
  bind_cols(predict(lda_fit, new_data = grid_vals))

grid_preds_qda <- grid_vals %>%
  bind_cols(predict(qda_fit, new_data = grid_vals))

grid_preds_log <- grid_vals %>%
  bind_cols(predict(log_fit, new_data = grid_vals))

plot_decision_boundary <- function(grid_preds, holdout_data, model_name) {
  ggplot() +
    geom_tile(data = grid_preds, aes(x = Red, y = Green, fill = .pred_class), alpha = 0.3) +
    geom_point(data = holdout_data, aes(x = Red, y = Green, color = Class), shape = 21, size = 2) +
    labs(
      title = paste("Decision Boundary on Test Set -", model_name),
      fill = "Predicted Class",
      color = "True Class"
    ) +
    theme_minimal()
}

plot_decision_boundary(grid_preds_lda, holdout_class, "LDA")
plot_decision_boundary(grid_preds_qda, holdout_class, "QDA")
plot_decision_boundary(grid_preds_log, holdout_class, "Logistic Regression")
```
```{r}
lda_preds_holdout <- predict(lda_fit, holdout_class) %>%
  bind_cols(holdout_class) %>%
  mutate(
    match = ifelse(.pred_class == Class, "Correct", "Incorrect")
  )

ggplot(lda_preds_holdout, aes(x = Red, y = Green)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.4, size = 0.8) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "LDA Predictions on Test Set",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()

```
```{r}
qda_preds_holdout <- predict(qda_fit, holdout_class) %>%
  bind_cols(holdout_class) %>%
  mutate(
    match = ifelse(.pred_class == Class, "Correct", "Incorrect")
  )

ggplot(qda_preds_holdout, aes(x = Red, y = Green)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.4, size = 0.8) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "QDA Predictions on Test Set",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()
```

```{r}
log_preds_holdout <- predict(log_fit, holdout_class) %>%
  bind_cols(holdout_class) %>%
  mutate(
    match = ifelse(.pred_class == Class, "Correct", "Incorrect")
  )

ggplot(log_preds_holdout, aes(x = Red, y = Green)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.4, size = 0.8) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "Logistic Regression Predictions on Test Set",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()
```

```{r}
library(dplyr)
library(ggplot2)


holdout_preds_lda <- holdout_class %>%
  bind_cols(predict(lda_fit, new_data = holdout_class))

holdout_preds_qda <- holdout_class %>%
  bind_cols(predict(qda_fit, new_data = holdout_class))

holdout_preds_log <- holdout_class %>%
  bind_cols(predict(log_fit, new_data = holdout_class))

plot_prediction_match <- function(pred_df, model_name) {
  pred_df <- pred_df %>%
    mutate(
      match = ifelse(Class == .pred_class, "Correct", "Incorrect")
    )
  
  ggplot(pred_df, aes(x = Red, y = Green)) +
    geom_point(aes(color = match, shape = Class), alpha = 0.5, size = 1.5) +
    scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
    labs(
      title = paste("Prediction Match on Test Set -", model_name),
      color = "Prediction Match",
      shape = "True Class"
    ) +
    theme_minimal()
}


plot_prediction_match(holdout_preds_lda, "LDA")
plot_prediction_match(holdout_preds_qda, "QDA")
plot_prediction_match(holdout_preds_log, "Logistic Regression")

```
```{r}
library(yardstick)
library(dplyr)


compute_recall <- function(pred_df, model_name) {
  pred_df <- pred_df %>%
    mutate(
      Class = factor(Class, levels = c("Non-Tarp", "Blue Tarp")),
      .pred_class = factor(.pred_class, levels = c("Non-Tarp", "Blue Tarp"))
    )
  
  recall_val <- recall(pred_df, truth = Class, estimate = .pred_class, event_level = "second")
  
  recall_val %>%
    mutate(model = model_name)
}


lda_preds <- holdout_class %>% bind_cols(predict(lda_fit, new_data = holdout_class))
qda_preds <- holdout_class %>% bind_cols(predict(qda_fit, new_data = holdout_class))
log_preds <- holdout_class %>% bind_cols(predict(log_fit, new_data = holdout_class))


recall_lda <- compute_recall(lda_preds, "LDA")
recall_qda <- compute_recall(qda_preds, "QDA")
recall_log <- compute_recall(log_preds, "Logistic Regression")


recall_results <- bind_rows(recall_lda, recall_qda, recall_log) %>%
  select(model, recall = .estimate)

recall_results

```
```{r}
recall_results %>%
  kable(
    caption = "Recall Values for LDA, QDA, and Logistic Regression on Test Set",
    digits = 3,
    col.names = c("Model", "Recall")
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
library(yardstick)
library(dplyr)


compute_confmat <- function(pred_df, model_name) {
  pred_df <- pred_df %>%
    mutate(
      Class = factor(Class, levels = c("Non-Tarp", "Blue Tarp")),
      .pred_class = factor(.pred_class, levels = c("Non-Tarp", "Blue Tarp"))
    )
  
  cm <- conf_mat(pred_df, truth = Class, estimate = .pred_class)
  
  list(model = model_name, cm = cm)
}


lda_preds <- holdout_class %>% bind_cols(predict(lda_fit, new_data = holdout_class))
qda_preds <- holdout_class %>% bind_cols(predict(qda_fit, new_data = holdout_class))
log_preds <- holdout_class %>% bind_cols(predict(log_fit, new_data = holdout_class))


lda_cm <- compute_confmat(lda_preds, "LDA")
qda_cm <- compute_confmat(qda_preds, "QDA")
log_cm <- compute_confmat(log_preds, "Logistic Regression")


lda_cm$cm
qda_cm$cm
log_cm$cm

```
```{r}
autoplot(lda_cm$cm, type = "heatmap") +
  labs(title = "Confusion Matrix - LDA (Test Set)")

autoplot(qda_cm$cm, type = "heatmap") +
  labs(title = "Confusion Matrix - QDA (Test Set)")

autoplot(log_cm$cm, type = "heatmap") +
  labs(title = "Confusion Matrix - Logistic Regression (Test Set)")
```
# SVM

```{r}
library(tidymodels)

svm_recipe <- recipe(Class ~ Red + Green + Blue, data = haititrain) %>%
  step_normalize(all_predictors())

svm_spec_rbf <- svm_rbf(mode = "classification") %>%
  set_engine("kernlab", prob.model = TRUE)


svm_wf_rbf <- workflow() %>%
  add_recipe(svm_recipe) %>%
  add_model(svm_spec_rbf)


svm_fit_rbf <- fit(svm_wf_rbf, data = haititrain)

svm_spec_linear <- svm_linear(mode = "classification") %>%
  set_engine("kernlab")

svm_wf_linear <- workflow() %>%
  add_recipe(svm_recipe) %>%
  add_model(svm_spec_linear)

svm_fit_linear <- fit(svm_wf_linear, data = haititrain)

svm_spec_poly <- svm_poly(mode = "classification") %>%
  set_engine("kernlab")

svm_wf_poly <- workflow() %>%
  add_recipe(svm_recipe) %>%
  add_model(svm_spec_poly)

svm_fit_poly <- fit(svm_wf_poly, data = haititrain)


```
```{r}
# RBF
svm_preds_rbf_train <- predict(svm_fit_rbf, haititrain, type = "prob") %>%
  bind_cols(predict(svm_fit_rbf, haititrain) %>% rename(.pred_class = .pred_class)) %>%
  bind_cols(haititrain %>% select(Class))

# Linear
svm_preds_linear_train <- predict(svm_fit_linear, haititrain, type = "prob") %>%
  bind_cols(haititrain %>% select(Class))

# Polynomial
svm_preds_poly_train <- predict(svm_fit_poly, haititrain, type = "prob") %>%
  bind_cols(haititrain %>% select(Class))

```

```{r}
svm_preds_rbf_train <- svm_preds_rbf_train %>%
  mutate(Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp")))

svm_preds_linear_train <- svm_preds_linear_train %>%
  mutate(Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp")))

svm_preds_poly_train <- svm_preds_poly_train %>%
  mutate(Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp")))
```

```{r}
library(yardstick)

svm_metrics <- metric_set(accuracy, sens, spec, precision, recall, f_meas, roc_auc)

# metrics for SVM RBF
svm_rbf_train_metrics <- svm_preds_rbf_train %>%
  mutate(.pred_class = if_else(`.pred_Blue Tarp` >= 0.5, "Blue Tarp", "Non-Tarp"),
         .pred_class = factor(.pred_class, levels = c("Blue Tarp", "Non-Tarp")),
         Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp"))) %>%
  svm_metrics(truth = Class, estimate = .pred_class, `.pred_Blue Tarp`, event_level = "first") %>%
  mutate(model = "SVM RBF")

# metrics for SVM Linear
svm_linear_train_metrics <- svm_preds_linear_train %>%
  mutate(.pred_class = if_else(`.pred_Blue Tarp` >= 0.5, "Blue Tarp", "Non-Tarp"),
         .pred_class = factor(.pred_class, levels = c("Blue Tarp", "Non-Tarp")),
         Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp"))) %>%
  svm_metrics(truth = Class, estimate = .pred_class, `.pred_Blue Tarp`, event_level = "first") %>%
  mutate(model = "SVM Linear")

# metrics for SVM Polynomial
svm_poly_train_metrics <- svm_preds_poly_train %>%
  mutate(.pred_class = if_else(`.pred_Blue Tarp` >= 0.5, "Blue Tarp", "Non-Tarp"),
         .pred_class = factor(.pred_class, levels = c("Blue Tarp", "Non-Tarp")),
         Class = factor(Class, levels = c("Blue Tarp", "Non-Tarp"))) %>%
  svm_metrics(truth = Class, estimate = .pred_class, `.pred_Blue Tarp`, event_level = "first") %>%
  mutate(model = "SVM Polynomial")
```

```{r}
library(tidyr)
library(dplyr)

svm_train_results <- bind_rows(
  svm_rbf_train_metrics,
  svm_linear_train_metrics,
  svm_poly_train_metrics
) %>%
  select(model, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate)

svm_train_results

```

```{r}
library(knitr)
library(kableExtra)

svm_train_results %>%
  kable(
    caption = "SVM Training Performance Metrics",
    digits = 4,
    format = "html"
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  )
```

```{r}
get_svm_metrics <- function(preds, model_name, threshold = 0.5) {
  preds <- preds %>%
    mutate(
      Class = factor(Class, levels = c("blue_tarp", "non_blue")),
      pred_class = if_else(`.pred_Blue Tarp` > threshold, "blue_tarp", "non_blue") %>% factor(levels = c("blue_tarp", "non_blue"))
    )

  tibble(
    model = model_name,
    accuracy = accuracy_vec(preds$Class, preds$pred_class),
    sensitivity = sens_vec(preds$Class, preds$pred_class, event_level = "first"),
    specificity = spec_vec(preds$Class, preds$pred_class, event_level = "first"),
    precision = precision_vec(preds$Class, preds$pred_class, event_level = "first"),
    recall = recall_vec(preds$Class, preds$pred_class, event_level = "first"),
    f_meas = f_meas_vec(preds$Class, preds$pred_class, event_level = "first"),
    roc_auc = roc_auc_vec(preds$Class, preds$`.pred_Blue Tarp`, event_level = "first")
  )
}

# Compute for each SVM model
svm_metrics_rbf <- get_svm_metrics(svm_preds_rbf, "SVM RBF")
svm_metrics_linear <- get_svm_metrics(linear_preds, "SVM Linear")
svm_metrics_poly <- get_svm_metrics(poly_preds, "SVM Polynomial")

# Combine results
svm_holdout_metrics <- bind_rows(svm_metrics_rbf, svm_metrics_linear, svm_metrics_poly)

# Display nicely
svm_holdout_metrics %>%
  kable(caption = "SVM Test Set Performance", digits = 4) %>%
  kable_styling(full_width = FALSE)


```

```{r}
library(dplyr)
library(yardstick)
library(ggplot2)

roc_data_train <- bind_rows(
  svm_preds_rbf_train %>% mutate(model = "SVM RBF"),
  svm_preds_linear_train %>% mutate(model = "SVM Linear"),
  svm_preds_poly_train %>% mutate(model = "SVM Polynomial")
)

roc_curve_train <- roc_data_train %>%
  group_by(model) %>%
  roc_curve(truth = Class, `.pred_Blue Tarp`, event_level = "first")

autoplot(roc_curve_train) +
  labs(
    title = "ROC Curves on Training Data",
    subtitle = "SVM RBF, Linear, and Polynomial Kernels",
    color = "Model"
  )

```

```{r}
roc_data_holdout <- bind_rows(
  svm_preds_rbf %>% mutate(model = "SVM RBF"),
  linear_preds %>% mutate(model = "SVM Linear"),
  poly_preds %>% mutate(model = "SVM Polynomial")
)

roc_data_holdout <- bind_rows(
  svm_preds_rbf %>% mutate(model = "SVM RBF"),
  linear_preds %>% mutate(model = "SVM Linear"),
  poly_preds %>% mutate(model = "SVM Polynomial")
) %>%
  mutate(Class = factor(Class, levels = c("blue_tarp", "non_blue")))  # Convert here

roc_curve_holdout <- roc_data_holdout %>%
  group_by(model) %>%
  roc_curve(truth = Class, `.pred_Blue Tarp`, event_level = "first")

autoplot(roc_curve_holdout) +
  labs(
    title = "ROC Curves on Test Data",
    subtitle = "SVM RBF, Linear, and Polynomial Kernels",
    color = "Model"
  )
```

```{r}
pca_overlay <- pca_pred_holdout %>%
  mutate(
    match = ifelse(Class == .pred_class, "Correct", "Incorrect")
  )

ggplot(pca_overlay, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.2) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "Test Set: True vs Predicted Class Overlay (SVM RBF)",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()
```

```{r}
pca_recipe_holdout <- recipe(Class ~ Red + Green + Blue, data = haititrain) %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 2) %>%
  prep()

pca_pred_linear_holdout <- bake(pca_recipe_holdout, new_data = svm_preds_linear_holdout)


svm_preds_linear_holdout <- predict(svm_fit_linear, holdout_class) %>%
  bind_cols(holdout_class)

pca_pred_linear_holdout <- bake(pca_recipe_holdout, new_data = svm_preds_linear_holdout)


pca_pred_linear_holdout$.pred_class <- svm_preds_linear_holdout$.pred_class

pca_overlay_linear <- pca_pred_linear_holdout %>%
  mutate(match = ifelse(Class == .pred_class, "Correct", "Incorrect"))

ggplot(pca_overlay_linear, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.2) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "Test Set: True vs Predicted Class Overlay (SVM Linear)",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()

```

```{r}
svm_preds_poly_holdout <- predict(svm_fit_poly, holdout_class, type = "class") %>%
  bind_cols(holdout_class) %>%
  rename(.pred_class = .pred_class)


pca_pred_poly_holdout <- bake(pca_recipe_holdout, new_data = svm_preds_poly_holdout)

# Add .pred_class manually
pca_pred_poly_holdout$.pred_class <- svm_preds_poly_holdout$.pred_class

pca_overlay_poly <- pca_pred_poly_holdout %>%
  mutate(match = ifelse(Class == .pred_class, "Correct", "Incorrect"))

ggplot(pca_overlay_poly, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = match, shape = Class), alpha = 0.2) +
  scale_color_manual(values = c("Correct" = "darkgreen", "Incorrect" = "red")) +
  labs(
    title = "Test Set: True vs Predicted Class Overlay (SVM Polynomial)",
    color = "Prediction Match",
    shape = "True Class"
  ) +
  theme_minimal()
```

```{r}
compute_confmat <- function(pred_df, model_name) {
  pred_df <- pred_df %>%
    mutate(
      Class = factor(Class, levels = c("Non-Tarp", "Blue Tarp")),
      .pred_class = factor(.pred_class, levels = c("Non-Tarp", "Blue Tarp"))
    )
  
  cm <- conf_mat(pred_df, truth = Class, estimate = .pred_class)
  
  list(model = model_name, cm = cm)
}

svm_preds_rbf_holdout <- holdout_class %>% bind_cols(predict(svm_fit_rbf, new_data = holdout_class))
svm_preds_linear_holdout <- holdout_class %>% bind_cols(predict(svm_fit_linear, new_data = holdout_class))
svm_preds_poly_holdout <- holdout_class %>% bind_cols(predict(svm_fit_poly, new_data = holdout_class))


svm_rbf_cm <- compute_confmat(svm_preds_rbf_holdout, "SVM RBF")
svm_linear_cm <- compute_confmat(svm_preds_linear_holdout, "SVM Linear")
svm_poly_cm <- compute_confmat(svm_preds_poly_holdout, "SVM Polynomial")


svm_rbf_cm$cm
svm_linear_cm$cm
svm_poly_cm$cm


autoplot(svm_rbf_cm$cm, type = "heatmap") + labs(title = "Confusion Matrix - SVM RBF (Test Set)")
autoplot(svm_linear_cm$cm, type = "heatmap") + labs(title = "Confusion Matrix - SVM Linear (Test Set)")
autoplot(svm_poly_cm$cm, type = "heatmap") + labs(title = "Confusion Matrix - SVM Polynomial (Test Set)")
```


# KNN

```{python}
import matplotlib.pyplot as plt
import numpy as np
blue_example=True

if blue_example:
    filename = 'orthovnir067_ROI_Blue_Tarps.txt'
else:
    filename = 'orthovnir067_ROI_NOT_Blue_Tarps.txt'

lines = open(filename).readlines()

lines = lines[8:]
lines[0]
xx = []
yy = []
rgbl = []
coord_x = 0
coord_y = 0

for line in lines:
    items = line.split()
    rgbl.append( (int(items[7])/256,int(items[8])/256,int(items[9])/256) )

rgbl = sorted(rgbl)

for point in rgbl:
    xx.append(coord_x)
    yy.append(coord_y)
    if coord_x == 100:
        coord_x = 0
        coord_y += 1
    else:
        coord_x += 1
        
# Define X and Y coordinates for the points
x = np.array(xx)
y = np.array(yy)

# Define RGB colors for each point (values must be between 0 and 1)
# Each inner tuple represents (Red, Green, Blue)
colors = np.array(rgbl)

# Create the scatter plot
plt.figure(figsize=(8, 6))
plt.scatter(x, y, c=colors,
            s=20)

# Add labels and title
plt.xlabel("X-axis")
plt.ylabel("Y-axis")
plt.title(filename)
plt.grid(True)

# Display the plot
plt.show()

```


```{python}
from PIL import Image

image_path = "orthovnir078_makeshift_villiage1.jpg"  # Replace with the actual path to your JPG file
with Image.open(image_path) as img:
    img = img.convert('RGB')  # Ensure the image is in RGB mode

width, height = img.size

rgb_values = []
for y in range(height):
    for x in range(width):
        r, g, b = img.getpixel((x, y))
        rgb_values.append((r, g, b))


#print(rgb_values)

with open('image_rgb.txt', 'w') as target:
    for rgb_value in rgb_values:
        target.write(f"{rgb_value[0]} {rgb_value[1]} {rgb_value[2]}\n")


lines = open('LOGthepreds.csv', 'r').readlines()

updated_rgb_values = []
for line,rgb in zip(lines[1:],rgb_values):
    if 'Non-Tarp' in line:
        updated_rgb_values.append((0,256,0))
    else:
        updated_rgb_values.append(rgb)


image = Image.new('RGB', (width, height))
image.putdata(updated_rgb_values)
image.save('LOGoutput_image.png')
```


```{r load-packages}
knitr::opts_chunk$set(cahce = TRUE, autodep = TRUE)
library(tidyverse)
library(tidymodels)
library(ggcorrplot)
library(GGally)
library(patchwork)
set.seed(6116)
```

```{r load-data}
train <- read_csv('data/HaitiPixels.csv') %>% 
  mutate(Class = factor(ifelse(Class == 'Blue Tarp', 'blue_tarp', 'non_blue')))

test <- read.table('holdout.txt')
```

```{r set-up-for-training}
resamples <- vfold_cv(train)

formula <- Class ~ Red + Green + Blue

knn_rec <- recipe(formula=formula, data=train) %>% 
  step_pca(all_numeric_predictors(), num_comp=parsnip::tune()) # saw a warning tune() may be masked
                                                               # so calling as parsnip::tune() to be sure

knn_model <- nearest_neighbor(neighbors=parsnip::tune()) %>%
  set_mode("classification") %>%
  set_engine("kknn")

knn_wf <- workflow() %>% 
  add_recipe(knn_rec) %>% 
  add_model(knn_model)

parameters <- extract_parameter_set_dials(knn_wf) %>% 
  update(
    num_comp=num_comp(c(1,4)),
    neighbors = neighbors(c(2, 50))
  )
```

```{r tune-knn, cache = TRUE}
knn_tune <- knn_wf %>% 
  tune_grid(
    resamples=resamples,
    grid=grid_space_filling(parameters, size = 10),
    control=control_resamples(save_pred = TRUE)
    )
```
```{r}
#| out.width: 75%
#| fig.width: 10
#| fig.height: 6
#| dev: "png"
autoplot(knn_tune, metric='roc_auc') +
  geom_line()
```


```{r knn-show-best}
show_best(knn_tune, metric = 'roc_auc')
```
The best parameter combination, both visually, and as per `show_best` is neighbors=34, `num_comp`=2. To prevent overfitting we can choose a simpler model, (less number of neighbors).  We can use `select_by_one_std_err` and see if there's a model withing one standard error that has a lower nuber of neighbors.
```{r visualize-bests}
knn_best <- knn_tune %>%
  select_by_one_std_err(neighbors, metric = 'roc_auc') %>% 
  dplyr::select(c(neighbors, num_comp))
knn_best
```
So it looks like 18 neighbors is within one str err and will perform similarly without so much danger of overfitting.


```{r}
knn_fit <- knn_wf %>% 
  finalize_workflow(knn_best) %>% 
  fit_resamples(
    resamples,
    control=control_resamples(save_pred = TRUE)
    )
```



```{r knn-roc-auc}
#| fig.width: 9
#| fig.height: 5
#| fig.align: center
#| out.width: 70%

#| dev: "png"
#| dpi: 100
# plot ROC AUC
knn_preds <-  collect_predictions(knn_fit)
knn_preds %>% 
  group_by(id) %>% 
  roc_curve(truth = Class, '.pred_blue_tarp', event_level = 'first') %>% 
  autoplot() +
  theme(legend.position = 'none')
  
```



```{r knn-cm}
conf_mat(knn_preds, truth = Class, estimate = .pred_class)
```

```{r knn-threshold}
knn_thresh_perf <- knn_preds %>%
  probably::threshold_perf(
    Class,
    .pred_blue_tarp,
    thresholds = seq(0.01, 0.99, 0.01),
    event_level = 'second',
    metrics = metric_set(
      j_index,
      yardstick::specificity,
      yardstick::sensitivity,
      accuracy
      )
    )
```



```{r knn-metric-plot}
max_sens <- knn_thresh_perf %>% 
  filter(.metric == 'sensitivity') %>% 
  filter(.estimate == max(.estimate))

max_j <- knn_thresh_perf %>% 
  filter(.metric == 'j_index') %>% 
  filter(.estimate == max(.estimate))
  
```




```{r knn-thresholds}
#| fig.width: 9
#| fig.height: 5
#| fig.align: center
#| out.width: 70%
#| fig.cap: ROC AUC based on nearest neighbors.
#| dev: "png"
#| dpi: 100
ggplot(knn_thresh_perf, aes(x = .threshold, y = .estimate, color = .metric)) +
  geom_line() +
  geom_vline(data = max_sens, aes(xintercept = .threshold, color = .metric)) +
  geom_vline(data = max_j, aes(xintercept = .threshold, color = .metric)) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = 'Threshold', y = '.metric')

```





```{r knn-conf-mtx}
knn_preds_threshold <- knn_preds %>% 
  mutate(.pred_class = factor(ifelse(.pred_blue_tarp>= 0.1, 'blue_tarp', 'non_blue')))

conf_mat(knn_preds_threshold, truth = Class, estimate = .pred_class)
```

Generating the confusion matrix with different thresholds it seems a value of 0.1 allows for only 30 of 2022 `blue_tarp` misclassifications.  That's about 1.5%.  A value of 0.2 threshold rises to about 2%.  In the present use case, it is more important to identify blue tarps than it is to miss them, so I will use 0.1 as the threshold

```{r knn-final}
# fit the final model with the entire training dataset
knn_final <- knn_wf %>% 
  finalize_workflow(knn_best) %>% 
  fit(train)
```


```{r knn-test-metrics}
get_metrics <- metric_set(
  j_index,
  yardstick::sensitivity,
  yardstick::specificity,
  accuracy)

# get metrics on training data
knn_train_metrics <- knn_preds_threshold %>% 
  get_metrics(truth = Class, estimate = .pred_class)

# get metrics on test data
knn_test_metrics <- parsnip::augment(knn_final, new_data = train) %>%
  get_metrics(truth = Class, estimate = .pred_class)

```




```{r knn-compare-metrics}
bind_rows(knn_train_metrics %>% mutate(data ='train'),
          knn_test_metrics %>%  mutate(data = 'test')) %>% 
  pivot_wider(id_cols =.metric,
              names_from = data,
              values_from = .estimate)
```



```{r}
knn_preds_threshold
```

```{r recall, message=FALSE}
library(caret)

cm <- confusionMatrix(data=knn_preds_threshold$.pred_class, reference=knn_preds_threshold$Class, positive="blue_tarp")
cm$byClass["F1"]
```



```{r set-up-for-training}
resamples <- vfold_cv(train)

formula <- Class ~ Red + Green + Blue

knn_rec <- recipe(formula=formula, data=train) %>% 
  step_pca(all_numeric_predictors(), num_comp=parsnip::tune()) # saw a warning tune() may be masked
                                                               # so calling as parsnip::tune() to be sure

knn_model <- nearest_neighbor(neighbors=parsnip::tune()) %>%
  set_mode("classification") %>%
  set_engine("kknn")

knn_wf <- workflow() %>% 
  add_recipe(knn_rec) %>% 
  add_model(knn_model)

parameters <- extract_parameter_set_dials(knn_wf) %>% 
  update(
    num_comp=num_comp(c(1,4)),
    neighbors = neighbors(c(2, 50))
  )
```

```{r tune-knn, cache = TRUE}
knn_tune <- knn_wf %>% 
  tune_grid(
    resamples=resamples,
    grid=grid_space_filling(parameters, size = 10),
    control=control_resamples(save_pred = TRUE)
    )
```

```{r visualize-bests}
knn_best <- knn_tune %>%
  select_by_one_std_err(neighbors, metric = 'roc_auc') %>% 
  dplyr::select(c(neighbors, num_comp))
knn_best
```

```{r}
knn_fit <- knn_wf %>% 
  finalize_workflow(knn_best) %>% 
  fit_resamples(
    resamples,
    control=control_resamples(save_pred = TRUE)
    )
```

```{r knn-roc-auc}
#| fig.width: 9
#| fig.height: 5
#| fig.align: center
#| out.width: 70%

#| dev: "png"
#| dpi: 100
# plot ROC AUC
knn_preds <-  collect_predictions(knn_fit)
knn_preds %>% 
  group_by(id) %>% 
  roc_curve(truth = Class, '.pred_blue_tarp', event_level = 'first') %>% 
  autoplot() +
  theme(legend.position = 'none')
  
```

```{r knn-conf-mtx}
knn_preds_threshold <- knn_preds %>% 
  mutate(.pred_class = factor(ifelse(.pred_blue_tarp >= 0.1, 'blue_tarp', 'non_blue')))
```

```{r knn-final}
# fit the final model with the entire training dataset
knn_final <- knn_wf %>% 
  finalize_workflow(knn_best) %>% 
  fit(train)
```

```{r}
test_preds_fmeas2 <- holdout %>%
  bind_cols(
    predict(knn_final, new_data = holdout, type = "class")) %>%
  mutate(Class = factor(Class))
```

```{r}
test_preds_fmeas2$color <- rgb(test_preds_fmeas2$Red, test_preds_fmeas2$Green, test_preds_fmeas2$Blue, maxColorValue = 255)

predictedcolorplot <- test_preds_fmeas2 %>%
  filter(.pred_class == "blue_tarp")

g3 <- ggplot(predictedcolorplot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "KNN\nTest Set Predicted Blue Tarp Colors")
g3
```

```{r}
incorrectcolorplot <- test_preds_fmeas2 %>%
  filter(.pred_class == "blue_tarp" & Class == "Non_Tarp")

g4 <- ggplot(incorrectcolorplot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "KNN\nTest Set Incorrectly Predicted Blue Tarp Colors")
g4
```





# BOOSTING


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(cache=FALSE, autodep=TRUE)
knitr::opts_chunk$set(fig.align="center", fig.pos="tbh")
set.seed(123)
library(future)
plan(multisession, workers = parallel::detectCores(logical = FALSE))
```



```{r}
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(tidymodels)
library(tidyverse)
library(GGally)
library(patchwork)
library(ggcorrplot)
library(dplyr)
library(lightgbm)
library(bonsai)
```


# Read in Training Dataset
```{r Read in Training Data}
#haiti <- read.csv('~/Documents/MSDS/DS6030/Disaster Relief Project/Data/HaitiPixels.csv')
setwd('~/Desktop/DS6030/')
haiti_train<- read.csv("HaitiPixels (1).csv")
haiti_train$Class <- ifelse(haiti_train$Class == "Blue Tarp", "Blue Tarp", "Not Blue Tarp")
haiti_resamples <- vfold_cv(haiti_train, strata=Class)
haiti_cv_control <- control_resamples(save_pred=TRUE)
haiti_cv_metrics <- metric_set(accuracy, roc_auc, recall, f_meas)
```


```{r holdout-load, message=FALSE}
process_load_holdout_file <- function(filename) {
    fileLines <- readLines(filename, skip=7)
    fileLines[8] <- gsub(";", "", fileLines[8])
    fileLines[8] <- gsub("Map X", "Map_X", fileLines[8])
    fileLines[8] <- gsub("Map Y", "Map_Y", fileLines[8])
    writeLines(fileLines, sprintf("temp_%s", filename))
    read_table(sprintf("temp_%s", filename), skip=7)
}

rename_set_target <- function(df, class_target) {
    colnames(df)[colnames(df) == 'B1'] <- 'Red'
    colnames(df)[colnames(df) == 'B2'] <- 'Green'
    colnames(df)[colnames(df) == 'B3'] <- 'Blue'
    df %>%mutate(target = class_target)
}
```

```{r holdout-2, message=FALSE}
setwd("~/Desktop/DS6030/HoldOutData")
# collect holdout blue tarp data
holdout_blue_tarp <- bind_rows(
    process_load_holdout_file('orthovnir067_ROI_Blue_Tarps.txt'),
    process_load_holdout_file('orthovnir069_ROI_Blue_Tarps.txt'),
    process_load_holdout_file('orthovnir078_ROI_Blue_Tarps.txt')
)

test_blue_tarp <- holdout_blue_tarp %>%
    rename_set_target('blue_tarp') %>%
    dplyr::select(c(Red, Green, Blue, target))

head(test_blue_tarp)
nrow(test_blue_tarp)
```
```{r holdout-3, message=FALSE}
# collect holdout non_blue tarp data
setwd("~/Desktop/DS6030/HoldOutData")
holdout_non_blue_tarp <- bind_rows(
  process_load_holdout_file('orthovnir057_ROI_NON_Blue_Tarps.txt'),
  process_load_holdout_file('orthovnir067_ROI_Blue_Tarps_data.txt'),
  process_load_holdout_file('orthovnir069_ROI_NOT_Blue_Tarps.txt'),
  process_load_holdout_file('orthovnir078_ROI_NON_Blue_Tarps.txt')
)

test_non_blue_tarp <- holdout_non_blue_tarp %>%
    rename_set_target('non_blue') %>%
    dplyr::select(c(Red, Green, Blue, target))

#head(holdout_non_blue_tarp)
nrow(test_non_blue_tarp)
```

```{r holdout-4, message=FALSE}
# combine both for final set
holdout <- bind_rows(
  test_blue_tarp,
  test_non_blue_tarp
)
head(holdout)
```

```{r}
write.table(x = holdout, file = "holdout.txt", sep = "\t", row.names = FALSE, quote=FALSE)
write.csv(holdout, 'holdout.csv', row.names=FALSE, quote=FALSE)
test <- read_table('holdout.txt')
nrow(test)
head(test)
```

```{r}
haiti_test <- test %>%
  rename(Class = target)
```
```{r}
haiti_test <- haiti_test %>%
  mutate(
    Class = ifelse(Class == "blue_tarp", "Blue Tarp", "Not Blue Tarp"),
    Class = factor(Class, levels = c("Not Blue Tarp", "Blue Tarp"))
  )

```

```{r}
head(haiti_test)
```

```{r}
haiti_formula <- (Class ~ Red + Green + Blue)
haiti_boost_wf <- workflow() %>%
add_recipe(recipe(haiti_formula, data=haiti_train)) %>%
add_model(boost_tree(mode="classification", engine="lightgbm",
trees=250, tree_depth=tune(), learn_rate=tune(), min_n = tune()))
```

```{r}
haiti_parameters <- extract_parameter_set_dials(haiti_boost_wf)
haiti_tune_boost <- tune_bayes(haiti_boost_wf,
resamples=haiti_resamples,
metrics=haiti_cv_metrics,
param_info=haiti_parameters, iter=25)
```
```{r}
autoplot(haiti_tune_boost)
```



```{r}
roc_cv_plot <- function(model_cv, model_name) {
  cv_predictions <- collect_predictions(model_cv)
  cv_roc <- cv_predictions %>%
    mutate(Class = factor(Class, levels = c("Not Blue Tarp", "Blue Tarp")))%>%
    roc_curve(truth = Class, `.pred_Blue Tarp`, event_level = "second")
return(autoplot(cv_roc) + labs(title=model_name))
}
```

```{r}
select_best(haiti_tune_boost, metric="roc_auc")
```

```{r}
best_haiti_boost_wf <- haiti_boost_wf %>%
finalize_workflow(select_best(haiti_tune_boost, metric="roc_auc"))
haiti_boost_cv <- fit_resamples(best_haiti_boost_wf, haiti_resamples, metrics=haiti_cv_metrics, control=haiti_cv_control)
g_roc_boost <- roc_cv_plot(haiti_boost_cv, "Boosting")
```
```{r}
g_roc_boost
```


```{r}
cv_metrics_summary <- collect_metrics(haiti_boost_cv)
cv_metrics_summary
```



```{r}
final_haiti_boost_model <- fit(best_haiti_boost_wf, data = haiti_train)
```
```{r}
test_preds <- augment(final_haiti_boost_model, new_data = haiti_test) %>%
  mutate(
    .pred_class = factor(.pred_class, levels = c("Not Blue Tarp", "Blue Tarp"))
  )

```



```{r}
test_acc <- accuracy(test_preds, truth = Class, estimate = .pred_class)
test_recall <- recall(test_preds, truth = Class, estimate = .pred_class, event_level = "second")
test_fmeas <- f_meas(test_preds, truth = Class, estimate = .pred_class, event_level = "second")
```

```{r}
test_roc<- roc_auc(test_preds, truth = Class, `.pred_Blue Tarp`, event_level = "second")
```
```{r}
test_metrics_summary <- bind_rows(test_acc, test_roc, test_fmeas, test_recall)
```

```{r}
cv_metrics_summary %>%
  select(.metric, mean) %>%
  rename(train = mean) %>%
  inner_join(
    test_metrics_summary %>%
      select(.metric, .estimate) %>%
      rename(test = .estimate),
    by = ".metric"
  )
```

# RANDOM FOREST

# Load and Prep Data
```{r setup-parallel}
#| cache: FALSE
#| message: false
library(future)
plan(multisession, workers = future::availableCores())
knitr::opts_chunk$set(cache = TRUE)
```

### Read in Training Data (HaitiPixels.csv)
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
```

```{r}
haititrain <- read.csv("https://gedeck.github.io/DS-6030/project/HaitiPixels.csv")
```

### Read in Holdout Data
```{r message=FALSE, warning=FALSE}
setwd("~/Documents/MSDS/DS6030/Disaster Relief Project/Data/HoldOutData")

files = sort(
c("orthovnir078_ROI_NON_Blue_Tarps.txt", "orthovnir078_ROI_Blue_Tarps.txt",
"orthovnir069_ROI_NOT_Blue_Tarps.txt", "orthovnir069_ROI_Blue_Tarps.txt",
"orthovnir067_ROI_NOT_Blue_Tarps.txt","orthovnir067_ROI_Blue_Tarps.txt",
"orthovnir057_ROI_NON_Blue_Tarps.txt"))

files2 = "orthovnir067_ROI_Blue_Tarps_data.txt"

col_names <- c("ID", "X", "Y", "Map_X", "Map_Y", "Lat", "Lon", "B1", "B2", "B3")

txt_data <- lapply(files, read.table, header = TRUE, fill = TRUE, skip = 7, col.names = col_names)

short_txt <- read.table(files2, header = TRUE, fill = TRUE, col.names = c("B1","B2","B3"))

short_data <- tibble(
ID = NA, X = NA, Y = NA, Map_X = NA, Map_Y = NA,
Lat = NA, Lon = NA, B1 = short_txt$B1, B2 = short_txt$B2, B3 = short_txt$B3)

names(txt_data) <- files

all_data <- c(txt_data, `orthovnir067_ROI_Blue_Tarps_data.txt` = list(short_data))

holdout <- bind_rows(all_data, .id = "source")
```

### Relabel Training & Holdout
```{r}
holdout <- holdout %>%
  rename(
    Red = B1,
    Green = B2,
    Blue = B3
  )
```

```{r}
haititrain <- haititrain %>%
  mutate(
    Class = if_else(Class == "Blue Tarp", "Blue_Tarp", "Non_Tarp"),
    Class = factor(Class)
  )

holdout <- holdout %>%
  mutate(Class = case_when(
    str_detect(source, "NOT|NON") ~ "Non_Tarp",
    TRUE ~ "Blue_Tarp"
  ))
```

### Split Data
```{r}
set.seed(123)
data_split <- initial_split(haititrain, strata = Class)
train_data <- training(data_split)
test_data  <- testing(data_split)
```

# Set up Recipes/Resamples/CV_Control/CV_metrics/Engine
```{r}
pixel_recipe <- recipe(Class ~ Red + Green + Blue, data = train_data) %>%
  step_normalize(all_predictors())
resamples <- vfold_cv(train_data, strata=Class)
cv_control <- control_resamples(save_pred=TRUE)
cv_metrics <- metric_set(roc_auc, accuracy, f_meas, recall)
```

```{r}
library(ranger)

rf <- rand_forest(mode = "classification", trees=tune(), min_n=tune(), mtry=tune()) %>%
  set_engine("ranger", importance = "impurity", probability = TRUE)
```



# 1st RF Model
### Set up Workflow
```{r}
library(bonsai)

rf_wf <- workflow() %>%
    add_recipe(pixel_recipe) %>%
    add_model(rf)

parameters <- extract_parameter_set_dials(rf_wf)
parameters
```

# Set Parameters / Tune
Tuning this model pretty wide.
```{r tune-1}
parameters <- extract_parameter_set_dials(rf_wf) %>%
    update(mtry = mtry(c(1, 3)))
tune_random <- tune_bayes(rf_wf,
    resamples=resamples,
    metrics=cv_metrics,
    param_info=parameters, iter=25)
autoplot(tune_random)
```

## Best Model (f_meas)
```{r}
final_params_fmeas <- select_best(tune_random, metric = "f_meas") 

final_params_fmeas %>%
  select(mtry, trees, min_n) %>%
  knitr::kable(caption = "Final Tuned Parameters for Random Forest Model") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

```{r}
show_best(tune_random, metric = "f_meas")
```

### Finalize Workflow (f_meas)
Finalize the workflow on best f_meas.
```{r bestfmeastraining}
#| warning: FALSE
best_rf_wf_fmeas <- rf_wf %>%
    finalize_workflow(select_best(tune_random, metric="f_meas"))
random_cv_fmeas <- fit_resamples(best_rf_wf_fmeas, resamples, metrics=cv_metrics, control=cv_control)
```

Fit the best f_meas on training data
```{r bestfmeasfit}
final_model_fmeas <- fit(best_rf_wf_fmeas, data = train_data)
```

Predict on the Holdout Set
```{r bestfmeastest}
test_preds_fmeas <- predict(final_model_fmeas, new_data = holdout, type = "prob") %>%
  bind_cols(predict(final_model_fmeas, new_data = holdout, type = "class")) %>%
  bind_cols(holdout %>% select(Class)) %>%
  mutate(Class = factor(Class))
```

Confusion Matrix for f_meas model
```{r}
conf_mat(test_preds_fmeas, truth = Class, estimate = .pred_class) %>%
  autoplot(type = 'heatmap')
```

F_meas metrics table
```{r}
train_metrics_fmeas <- collect_metrics(random_cv_fmeas) %>%
  mutate(model = "Train") %>%
  select(.metric, model, .estimate = mean)

test_metrics_fmeas <- bind_rows(
  accuracy(test_preds_fmeas, truth = Class, estimate = .pred_class),
  f_meas(test_preds_fmeas, truth = Class, estimate = .pred_class),
  recall(test_preds_fmeas, truth = Class, estimate = .pred_class),
  roc_auc(test_preds_fmeas, truth = Class, `.pred_Blue_Tarp`)) %>%
  mutate(model = "Test") %>%
  select(.metric, model, .estimate)

combined_metrics_fmeas <- bind_rows(train_metrics_fmeas, test_metrics_fmeas) %>%
  pivot_wider(names_from = model, values_from = .estimate) %>%
  rename(Metric = .metric,
    `Train Estimate` = Train,
    `Test Estimate` = Test)

knitr::kable(combined_metrics_fmeas, digits = 4, caption = "Random Forest on Training and Test Sets") %>%
  kableExtra::kable_styling(full_width = FALSE)
```


# 2nd RF Model
### Set Parameters / Tune
```{r}
parameters <- extract_parameter_set_dials(rf_wf) %>%
    update(mtry = mtry(c(1, 3)),
           trees = trees(c(0, 700)),
           min_n = min_n(c(10, 40)))
second_tune <- tune_bayes(rf_wf,
    resamples=resamples,
    metrics=cv_metrics,
    param_info=parameters, iter=25)
autoplot(second_tune)
```

## Best Model (f_meas)
```{r}
final_params_fmeas2 <- select_best(second_tune, metric = "f_meas") 

final_params_fmeas2 %>%
  select(mtry, trees, min_n) %>%
  knitr::kable(caption = "Final Tuned Parameters for Random Forest Model") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

```{r}
show_best(second_tune, metric = "f_meas")
```

### Finalize Workflow (f_meas)
Finalize the workflow on best f_meas.
```{r}
#| warning: FALSE
best_rf_wf_fmeas2 <- rf_wf %>%
    finalize_workflow(select_best(second_tune, metric="f_meas"))
random_cv_fmeas2 <- fit_resamples(best_rf_wf_fmeas2, resamples, metrics=cv_metrics, control=cv_control)
```

Fit the best f_meas on training data
```{r}
final_model_fmeas2 <- fit(best_rf_wf_fmeas2, data = train_data)
```

Predict on the Holdout Set
```{r}
test_preds_fmeas2 <- holdout %>%
  bind_cols(
    predict(final_model_fmeas2, new_data = holdout, type = "prob"),
    predict(final_model_fmeas2, new_data = holdout, type = "class")) %>%
  mutate(Class = factor(Class))
```

Confusion Matrix for f_meas model
```{r}
conf_mat(test_preds_fmeas2, truth = Class, estimate = .pred_class) %>%
  autoplot(type = 'heatmap')
```

F_meas metrics table
```{r}
train_metrics_fmeas2 <- collect_metrics(random_cv_fmeas2) %>%
  mutate(model = "Train") %>%
  select(.metric, model, .estimate = mean)

test_metrics_fmeas2 <- bind_rows(
  accuracy(test_preds_fmeas2, truth = Class, estimate = .pred_class),
  f_meas(test_preds_fmeas2, truth = Class, estimate = .pred_class),
  recall(test_preds_fmeas2, truth = Class, estimate = .pred_class),
  roc_auc(test_preds_fmeas2, truth = Class, `.pred_Blue_Tarp`)) %>%
  mutate(model = "Test") %>%
  select(.metric, model, .estimate)

combined_metrics_fmeas2 <- bind_rows(train_metrics_fmeas2, test_metrics_fmeas2) %>%
  pivot_wider(names_from = model, values_from = .estimate) %>%
  rename(Metric = .metric,
    `Train Estimate` = Train,
    `Test Estimate` = Test)

knitr::kable(combined_metrics_fmeas2, digits = 4, caption = "Random Forest on Training and Test Sets") %>%
  kableExtra::kable_styling(full_width = FALSE)
```


```{r}
holdout$color <- rgb(holdout$Red, holdout$Green, holdout$Blue, maxColorValue = 255)

fullcolorplot <- holdout %>% 
  filter(Class == "Blue_Tarp")

g1 <- ggplot(holdout, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "Test Set Colors")
```

```{r}
g2 <- ggplot(fullcolorplot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "Test Set Blue Tarp Colors")
```

```{r fig.height=8, fig.width=15}
library(patchwork)
(g1 + g2)
```

```{r}
test_preds_fmeas2$color <- rgb(test_preds_fmeas2$Red, test_preds_fmeas2$Green, test_preds_fmeas2$Blue, maxColorValue = 255)

predictedcolorplot <- test_preds_fmeas2 %>%
  filter(.pred_class == "Blue_Tarp")

g3 <- ggplot(predictedcolorplot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "Random Forest\nTest Set Predicted Blue Tarp Colors")
g3
```

```{r}
incorrectcolorplot <- test_preds_fmeas2 %>%
  filter(.pred_class == "Blue_Tarp" & Class == "Non_Tarp")

g4 <- ggplot(incorrectcolorplot, aes(x = Red, y = Green)) +
  geom_point(aes(color = color), size = 2, alpha = 0.25) +
  scale_color_identity() +
  theme_minimal()+
  labs(title = "Random Forest\nTest Set Incorrectly Predicted Blue Tarp Colors")
g4
```
```{r}
setwd("~/Documents/MSDS/DS6030/Disaster Relief Project/Data/HoldOutData")
new_datapoints <- read_table('image_rgb.txt')
colnames(new_datapoints) <- c("Red", "Green", "Blue")
thepreds <- predict(final_model_fmeas2, new_data=new_datapoints)
write.csv(thepreds, "RFthepreds.csv", row.names = FALSE)
```

