---
title: "Home Sales Analysis in Kings County, Washington"
author: "Arjun Gautam"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center"
)

```

# Section 1: Summary of Findings

King County, Washington, home to Seattle, is known for its rich arts scene, scenic waterfront, and major attractions like the Seattle Art Museum and the original Starbucks coffee shop[^1]. It also serves as the headquarters for major companies such as Amazon, Boeing, and Microsoft, and is home to professional sports teams including the Seahawks (NFL), Kraken (NHL), and Mariners (MLB). With a population of over 2.2 million residents[^2], the county offers a variety of activities, strong economic opportunities, and outdoor attractions that draw residents from across the country.

[^1]: <https://visitseattle.org/things-to-do/sightseeing/top-25-attractions/>

[^2]: <https://kingcounty.gov/en/dept/executive/governance-leadership/performance-strategy-budget/regional-planning/demographics>

Given its advantages, King County is a competitive market for prospective home buyers. However, buyers must weigh a variety of factors when searching for the perfect home, including location, size, number of bedrooms and bathrooms, condition, year built, and more. All of these factors can significantly impact affordability, availability, and condition of their new home. In this report, we analyze the Kings County housing market in detail, highlighting the key factors that influence home prices and overall housing conditions.

Our team obtained a data set containing 21,613 house sale prices for King County Washington, from May 2014 through May 2015. The goal of our analysis was to explore how home prices relate to a variety of property characteristics, and to identify which factors are associated with higher-quality homes. We conducted our analysis by creating visualizations and fitting both a linear regression model and a logistic regression model to better understand how these characteristics influence home prices and conditions.

For both models, we used contextual reasoning to eliminate certain variables. For example, we excluded latitude and longitude due to their small range, and removed variables measuring the size of neighboring homes (`sqft_living15` and `sqft_lot15`), as our focus was on the characteristics of the home itself. After finalizing the full model, we applied various model selection techniques to improve model accuracy.

Our final multiple linear regression model predicts price using eleven predictors: `bedrooms`, `bathrooms`, `sqft_living`, `floors`, `waterfront`, `view`, `condition`, `grade`, `yr_built`, `Warm_Cold` (season), and `Metro_Area`. Based on our estimates, waterfront presence has the largest impact on price, followed by grade and Metro_Area. In the logistic regression model predicting whether a home is of good quality, being within city limits and the presence of a basement were the strongest predictors, with `sqft_living` also contributing significantly to the likelihood of high quality.

Overall, buyers seeking the luxuries of waterfront views, high-grade construction, and city living should expect to pay a premium. Similarly, prospective buyers can have greater confidence in the quality of a home if it is located within the city, includes a basement, and offers larger living space.

# Section 2: Data and the Variables

The original data set included 21,613 observations across 21 variables. The variables and their definitions are listed below:

```{r, echo = FALSE}

variables <- data.frame(
  Variable = c(
    "id", "date", "price", "bedrooms", "bathrooms", "sqft_living", 
    "sqft_lot", "floors", "waterfront", "view", "condition", "grade", 
    "sqft_above", "sqft_basement", "yr_built", "yr_renovated", 
    "zipcode", "lat", "long", "sqft_living15", "sqft_lot15"
  ),
  Description = c(
    "Unique identifier for each home sold",
    "Date of home sale, ranging from May 2014 through May 2015",
    "Sale price of the home, ranging from $75,000 to $7,700,000",
    "Number of bedrooms",
    "Number of bathrooms, where 0.5 represents a half bathroom with a toilet but no shower",
    "Square footage of the home's interior living space",
    "Square footage of the home's land space",
    "Number of floors",
    "Binary variable indicating whether the home overlooks the waterfront",
    "Numerical rating of the home's view quality, ranging from 0 to 4",
    "Numerical rating of the home's condition, ranging from 1 to 5",
    "Numerical rating of construction and design quality, ranging from 1–13 (1–3 = low, 7 = average, 11–13 = high)",
    "Square footage of interior living space above ground level",
    "Square footage of interior living space below ground level",
    "Year the home was initially built, ranging from 1900 to 2015",
    "Year the home was last renovated (0 indicates no renovation)",
    "Zip code of the home's location",
    "Latitude coordinate of the home, ranging from 47.16 to 47.78",
    "Longitude coordinate of the home, ranging from -122.5 to -121.3",
    "Average square footage of interior living space for the nearest 15 neighbors",
    "Average square footage of land space for the nearest 15 neighbors"
  )
)

# Print the table
knitr::kable(variables)
```

To better investigate the characteristics associated with home price and condition, we performed several variable transformations and creations. First, we simplified the `date` variable by removing the arbitrary time component. We also created a `season` variable based on the sale date, and due to the results of our regression, we created a `Warm_Cold` variable to group together warm and cold seasons. Because the `yr_renovated` variable labeled non-renovated homes as 0, we replaced it with a binary indicator for whether a home had been renovated. Similarly, we created a binary variable indicating whether a home had a basement, based on `sqft_basement`. We created an additional variable `recent.build`, representing either the original build year or the renovation year, whichever was more recent. We also created a binary indicator for whether a home was located within Seattle city limits based on `zipcode`, and a variable for the total number of rooms in a house. Finally, we created a `good_quality` indicator for homes with a `condition` greater than 3 and a `grade` greater than 7.

The variables we created and their definitions are listed below:

```{r, echo = FALSE}

variables2 <- data.frame(
  Variable = c(
    "season","renovated","basement","recent.build",
    "Metro_Area","good_quality", "total_rooms","Warm_Cold"
  ),
  Description = c(
    "Category variable indicating the season of the home sale, including 'Fall', 'Winter', 'Spring', 'Summer'",
    "Binary variable indicating whether the home has been renovated",
    "Binary variable indicating whether the home includes a basement",
    "Date of either original build year or renovation year, whichever is more recent",
    "Binary variable indicating whether the home is within Seattle city limits",
    "Binary variable indicating whether the home is good quality (condition > 3 and grade > 7)",
    "Total rooms in a home, found using the sum of bathrooms and bedrooms",
    "Binary variable indicating whether the home is sold during warm seasons (Spring, Summer) or cold seasons (Fall, Winter)"
  )
)

# Print the table
knitr::kable(variables2)
```

# Section 3: Data Issues and Resolutions

Our team identified three types of data entry errors. First, the observation with an `id` of 2402100895 listed 33 bedrooms, which was clearly unreasonable. Our team assumed that this was intended to be a 3 bedroom home, so we compare this home's characteristics and price to other 3 bedroom homes and determined it was likely a data entry error. As such, we corrected the number of bedrooms from 33 to 3.

Next, we found several observations with 0 bedrooms and multiple bathrooms, or multiple bedrooms with 0 or only half bathrooms. Our team argued that these do not represent livable residential properties, which our analysis focuses on, and decided to remove all observations that had either 0 bedrooms or 0 bathrooms.

In addition, the `date` variable contained unnecessary information, including time stamps that were arbitrary, as all showed 'T000000'. We removed the time component but retained the date. Since the raw dates were not helpful for our regression analysis, we created two new variables: `season`, indicating the season in which the home was sold, and `Warm_Cold`, further grouping the seasons.

Beyond individual observations, we also addressed issues with several variable issues. The `yr_renovated` variable was misleading, as homes that were never renovated were coded as 0. We replaced this with a binary `renovated` variable indicating whether a home had been renovated. Additionally, the variables `sqft_above` and `sqft_basement` were linearly related to `sqft_living`, potentially causing multicollinearity. To resolve this while showing the relevance of a basement, we removed both variables and created a binary `basement` variable, indicating whether a home has a basement. Finally, since `zipcodes` was not helpful for our analysis in its current form, we created a new binary variable, `Metro_Area`, based on whether a property's zipcode began with 981, indicating it was located within Seattle city limits.

# Section 4: Visualization of Price

```{r message=FALSE, warning=FALSE, echo = FALSE}
library(tidyverse)
library(datasets)
library(MASS)
library(ggplot2)
library(leaps) 
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
Data <-read.csv("kc_house_data.csv", header=TRUE)
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
## Reformat date
Data$date <- Data$date %>%
  substr(1,8) %>%
  as.Date(format = "%Y%m%d")

# Create variable for seasons
Data <- Data %>%
  mutate(
    month = as.numeric(format(date, "%m")),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    )
  )

Data <- Data %>%
  mutate(
    Warm_Cold = case_when(
      season == "Spring" | season == "Summer" ~ 1,
      season == "Fall"   | season == "Winter" ~ 0
    ),
    Warm_Cold = factor(Warm_Cold, levels = c(0,1))
  )

## Change record with 33 bedrooms to 3 (mistype)
Data[Data$id == 2402100895, "bedrooms"] <- 3

## Create binary variable for whether renovated
Data <- Data %>%
  mutate(renovated = if_else(yr_renovated == 0, 0, 1))

## Create binary variable for whether basement
Data <- Data %>%
  mutate(basement = if_else(sqft_basement == 0, 0, 1))

## Create variable for most recent year built
Data$recent.build <- pmax(Data$yr_built, Data$yr_renovated)

## Identify zipcodes within city limits
Data$zip_sub <- substr(Data$zipcode,2, 3) 
Data$Metro_Area <- ifelse(Data$zip_sub == '81', 1, 0)

## Create variable for time series
Data$days <- as.integer(Data$date - min(Data$date))

# Binary if good quality
Data$good_quality <- ifelse(Data$condition > 3 & Data$grade > 7, 1, 0)

## Change waterfront/zipcode to factors
Data$season <- factor(Data$season)
Data$waterfront <- factor(Data$waterfront, levels = c(0,1))
Data$zipcode <- factor(Data$zipcode)
Data$renovated <- factor(Data$renovated)
Data$Metro_Area <- factor(Data$Metro_Area)
Data$good_quality <- factor(Data$good_quality)
Data$basement <- factor(Data$basement)

## Filtering to make sure a house has a bedroom, bathroom, and lot/living area greater than 0
Data <- subset(Data, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]
```

In this section, we analyze the available predictors of home price using visualizations. To better understand our binary variables, we first created univariate visualizations showing the proportions of homes that are waterfront properties, renovated, sold during the warm seasons, or located within city limits. As shown below, only a small portion of the homes in our dataset are waterfront properties or have been renovated. However, a majority of homes were sold during the warmer spring and summer months, and just under half are located within Seattle city limits.

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
library(scales) 
library(ggthemes)

bardata <- train %>%
  dplyr::select(waterfront, renovated, Warm_Cold, Metro_Area) %>%
  pivot_longer(cols = everything(), names_to  = "variable", values_to = "value")

ggplot(bardata, aes(x = variable, fill = factor(value))) +
  geom_bar(position = "fill", alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Value") +
  labs(title = "Proportion for Each Binary Variable", x = "Variable", y = "Proportion") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
  axis.title = element_text(size = 10),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 8),
  legend.key.size = unit(0.5, "cm"))
```

Next, we visualized the distribution of home sale prices. Note that a majority of home sale prices are between 0 and 1 million, with the mean home sale price being near \$500,000. Note that there are some outlier prices, with prices reaching over 7 million dollars.

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
ggplot(train, aes(x = price))+
  geom_density(color = 'navyblue',fill = 'darkorange', alpha = 0.8, linewidth = 0.8)+
  theme_fivethirtyeight()+
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  labs(y = "Density", title = "Distribution of Home Sale Prices")+
  theme(plot.title = element_text(hjust = 0.5, size = 12),
  axis.title = element_text(size = 10))
```

To further investigate how the binary variables affected our sales prices, we introduced our binary variables to the density plots. While there do not seem to be stark differences between the two groups for some plots, we can see that waterfront severely affects the distribution, and renovations appear to slightly increase home value. This indicates that homes with waterfronts and renovations tend to be more expensive.

```{r echo=FALSE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
library(ggthemes)
library(gridExtra)

densitytheme <- theme(plot.title = element_text(hjust = 0.5, size = 10),
  axis.title = element_text(size = 8),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9),
  legend.key.size = unit(0.2, "cm"))

denswf <- ggplot(train, aes(x = price, color = waterfront, fill = waterfront)) +
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Waterfront?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Waterfront?") +
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  labs(x = "Price", y = "Density", title = "Price by Waterfront")+
  densitytheme

densr <- ggplot(train, aes(x = price, color = renovated, fill = renovated)) +
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Renovated?") +
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Renovated?") +
  labs(x = "Price", y = "Density", title = "Price by Renovated")+
  densitytheme

denswc <- ggplot(train, aes(x = price, color = Warm_Cold, fill = Warm_Cold)) +
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Warm Seasons?") +
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Warm Seasons?") +
  labs(x = "Price", y = "Density", title = "Price by Warm/Cold")+
  densitytheme

densma <- ggplot(train, aes(x = price, color = Metro_Area, fill = Metro_Area)) +
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Metro Area?") +
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Metro Area?") +
  labs(x = "Price", y = "Density", title = "Price by Metro Area")+
  densitytheme

grid.arrange(denswc, densma, densr, denswf, ncol = 2, nrow = 2)
```

The gap between waterfront properties and non-waterfront properties is even more evident in the following visualization, where we observe the mean prices of home sales for each value of our binary variables. Note that our previous claim about renovations is also affirmed here.

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
meandata <- train %>%
  pivot_longer(cols = c(renovated, waterfront, Warm_Cold, Metro_Area), names_to  = "variable", values_to = "value") %>%
  group_by(variable, value) %>%
  summarize(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

ggplot(meandata, aes(x = variable, y = mean_price, fill = factor(value))) +
  geom_bar(stat='identity', position = position_dodge(width = 0.8), width = 0.7, alpha = 0.8) +
  geom_text(aes(label = dollar(round(mean_price))), position = position_dodge(width = 0.8), 
    vjust = -0.5, size = 3) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M")) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("0" = "No", "1" = "Yes"), name = "Value") +
  labs(x = NULL, y = "Mean Price (millions)", title = "Mean Price by Variable") +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(hjust = 1), axis.title.y = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 9), legend.key.size = unit(0.2, "cm"))
```

We used the same strategy to view how condition and grade affect price. Both condition and grade have a significant direct relationship with price, with grade having a very steep increase for values greater than 7.

```{r echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE}
meandata2 <- train %>%
  group_by(condition) %>%
  summarize(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

meandata3 <- train %>%
  group_by(grade) %>%
  summarize(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

meanplotc <- ggplot(meandata2, aes(x = condition, y = mean_price, fill = factor(condition))) +
  geom_bar(stat = 'identity', width = 0.8, alpha = 0.8, fill = 'navyblue') +
  geom_text(aes(label = dollar(round(mean_price))), position = position_dodge(width = 0.8), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M")) +
  labs(x = NULL, y = "Mean Price (millions)", title = "Mean Price by Condition") +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(hjust = 1), axis.title.y = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 9), legend.key.size = unit(0.2, "cm"))

meanplotg <- ggplot(meandata3, aes(x = grade, y = mean_price, fill = factor(grade))) +
  geom_bar(stat = 'identity', width = 0.8, alpha = 0.8, fill = 'navyblue') +
  geom_text(aes(label = dollar(round(mean_price))), position = position_dodge(width = 0.8), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M")) +
  labs(x = NULL, y = "Mean Price (millions)", title = "Mean Price by Grade") +
  theme_fivethirtyeight() +
  theme(plot.title = element_text(hjust = 0.5, size = 12), axis.text.x = element_text(hjust = 1), axis.title.y = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 9), legend.key.size = unit(0.2, "cm"))

grid.arrange(meanplotc, meanplotg)
```

Finally, we used scatterplots to visualize how some of our quantitative predictors affect price. Note that bedrooms, bathrooms, and grade all appear to increase with price. However, condition does not appear to have a huge impact on price.

```{r echo=FALSE, fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
bedplot <- ggplot(train, aes(x = bedrooms, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange') +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  labs(x = "Number of Bedrooms", title = "Price by Number of Bedrooms") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

bathplot <- ggplot(train, aes(x = bathrooms, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange') +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  labs(x = "Number of Bathrooms", title = "Price by Number of Bathrooms") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

conditionplot <- ggplot(train, aes(x = condition, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange') +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  labs(x = "Condition", title = "Price by Condition") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

gradeplot <- ggplot(train, aes(x = grade, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange') +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  labs(x = "Grade", title = "Price by Grade") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

grid.arrange(bedplot, bathplot, conditionplot, gradeplot, ncol = 2, nrow = 2)
```

Diving even further, we visualized bedrooms and bathrooms affect on price, separated by homes with a waterfront and homes without. There is a very clear positive relationship between bedrooms, bathrooms and price, which becomes even more significant in homes with waterfronts.

```{r fig.height=10, fig.width=7, message=FALSE, warning=FALSE, echo = FALSE}
bedplot2 <- ggplot(train, aes(x = bedrooms, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange', se = FALSE) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  facet_wrap(train$waterfront)+
  labs(x = "Number of Bedrooms", title = "Price by Number of Bedrooms by Waterfront") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

bathplot2 <- ggplot(train, aes(x = bathrooms, y = price)) +
  geom_point(color = 'navyblue', alpha = 0.5) +
  geom_smooth(method = "lm", color = 'darkorange', se = FALSE) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)") +
  theme_fivethirtyeight() +
  facet_wrap(train$waterfront)+
  labs(x = "Number of Bathrooms", title = "Price by Number of Bathrooms by Waterfront") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10))

grid.arrange(bedplot2, bathplot2)
```

# Section 5: Linear Regression

```{r,warning=FALSE,message=FALSE, echo=FALSE}
library(tidyverse)
library(datasets)
library(MASS)
library(ggplot2)
library(leaps) 
library(faraway)
```

```{r, echo=FALSE}
Data <-read.csv("kc_house_data.csv", header=TRUE)
```

```{r, echo=FALSE}
## reformat date
Data$date <- Data$date %>%
  substr(1,8) %>%
  as.Date(format = "%Y%m%d")

Data <- Data %>%
  mutate(
    month = as.numeric(format(date, "%m")),
    season = case_when(
      month == 12 | month == 1 | month == 2  ~ "Winter",
      month == 3  | month == 4 | month == 5  ~ "Spring",
      month == 6  | month == 7 | month == 8  ~ "Summer",
      month == 9  | month == 10 | month == 11 ~ "Fall"
    )
  )

## factors
Data$waterfront <- factor(Data$waterfront, levels = c(0,1))
Data$zipcode <- factor(Data$zipcode)

## change 33 bedrooms to 3
Data[Data$id == 2402100895, "bedrooms"] <- 3

## renovated/not renovated
Data <- Data %>%
  mutate(renovated = if_else(yr_renovated == 0, 0, 1))

## most recent year of build
Data$recent.build <- pmax(Data$yr_built, Data$yr_renovated)

## Define metropolitan ZIP codes
Data$zip_sub <- substr(Data$zipcode,2, 3) 
Data$Metro_Area <- ifelse(Data$zip_sub == '81', 1, 0)
Data$Metro_Area <-factor(Data$Metro_Area)
Data$renovated <-factor(Data$renovated)
## filtering to make sure a house has a bedroom, bathroom, and lot/living area greater than 0
Data <- subset(Data, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)
```

#### Model 1: Full Regression

```{r, echo=FALSE}
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]

##intercept only model
regnull <- lm(price~1, data=train)
##model with all predictors
regfull <- lm(price~bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + condition + grade + yr_built + season + renovated + recent.build + Metro_Area, data=train)
summary(regfull)
```

To begin the process of fitting a linear regression, we decided to start with a full model of `price ~ bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + condition + grade + yr_built + season + renovated + recent.build + Metro_Area`. We decided to not include latitude and longitude because it didn't make sense contextually that a house with a higher or lower latitude or longitude would have a significant impact on price of houses, especially within such a small range. In addition, we excluded `sqft_above` and `sqft_basement` because they both affect the other predictor `sqft_living`. Lastly, we excluded `sqft_living15` and `sqft_lot15` because we wanted to focus on the variables that focused on the house itself and not the 15 closest houses in proximity to a house.

After fitting a multiple linear regression model to predict housing prices, we observed that the season variable had differences in significance between the different levels. Specifically, only the Spring season was a significant predictor of price, while Summer and Winter were not statistically significant with Fall as a reference variable. We decided to research more into the effect of seasons on housing prices, and our research showed that housing prices tend to be higher during the Spring and Summer months, while prices are generally lower during Fall and Winter. Based on these findings, we decided to transform the season variable into a new binary factor grouping Spring and Summer together as "Warm Months", and Fall and Winter as "Cold Months", with "Cold Months" being the reference class. This transformation is intended to better capture the broader seasonal trends affecting housing prices and to simplify the model. After this transformation, we refit the model to include the variable `Warm_Cold` instead of `Seasons`.

#### Model 2: Tranforming Season predictor to Warm/Cold Seasons

```{r, echo=FALSE}
Data <- Data %>%
  mutate(
    Warm_Cold = case_when(
      season == "Spring" | season == "Summer" ~ 1,
      season == "Fall"   | season == "Winter" ~ 0
    ),
    Warm_Cold = factor(Warm_Cold, levels = c(0,1))
  )
```

```{r, echo = FALSE}
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]
##model with all predictors
regfull2 <- lm(price~bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + condition + grade + yr_built  + Warm_Cold + renovated + recent.build + Metro_Area, data=train)
summary(regfull2)
```

After refitting the model using the new `Warm_Cold` variable, the `Warm_Cold` variable was shown to be significant in predicting the price of houses in Kind County. The estimated difference in the price between houses sold during Spring/Summer and Fall/Winter is \$16,260. Although the R-squared decreased slightly, the model became more simple and easier to interpret In this version of the model, all predictors are shown to be significant. However, we still want to check the predictors for multicollinearity.

##### Checking for Multicollinearity

```{r, echo=FALSE}
print(round(faraway::vif(regfull2),3))
```

To check for potential multicollinearity, we calculated the Variance Inflation Factors (VIFs) for all predictors in the model. Most variables showed VIF values below 5, however `yr_built` and `recent.build showed` very high VIFs (34.6 and 34.2 respectively). This suggests that there may be a strong linear relationship between the two variables. To address this issue, we decided to remove the `recent.build` variable and refit the model. By excluding it, we aim to reduce multicollinearity.

#### Model 3: Removing recent.build

```{r, echo=FALSE}
regfull3 <- lm(price~bedrooms + bathrooms + sqft_living + sqft_lot + floors + waterfront + view + condition + grade + yr_built + Warm_Cold + renovated + Metro_Area, data=train)

summary(regfull3)
```

##### Multicollinearity Check

```{r, echo=FALSE}
print(round(faraway::vif(regfull3),3))
```

After removing recent.build and refitting the model, we observe that all the remaining predictors are significant in predicting the price of houses in King County. In addition, when calculating the VIF values of all the predictors again, all the VIF values are less than 5, which tells us that multicollinearity is no longer a major concern and the predictors are relatively independent of each other. Now we will check if the regression assumptions are met for this linear regression or if transformations need to be made.

##### Checking for Assumptions

##### Diagnostic Plots

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
plot(regfull3)
```

##### VIFs

```{r, echo = FALSE}
print(round(faraway::vif(regfull3),3))
```

##### Box Cox Plot

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE, echo = FALSE}
library(MASS)
MASS::boxcox(regfull3)
```

In residual plot, we see the vertical spread of the residuals increase as we move from left to right, so assumption 2, "The errors, for each fixed value of x, have constant variance", is not met. However, the red line overlaid is along the x-axis with a slight curvature from this view. Now that we know that both assumptions 2, and possibly assumption 1, is not met, we need to transform the response variable first, to stabilize the variance. In the Box Cox plot, CI includes a number very close to 0, so we can choose $\lambda = 0$ to log transform the response variable to get $y^* = \log{(y)}$.

#### Model 4: Log Transformation

```{r, echo=FALSE}
train$log_price <- log(train$price)
reg_log <- lm(log_price ~bedrooms + bathrooms + sqft_living +sqft_lot + floors + waterfront + view + condition + grade + yr_built + Warm_Cold + renovated + Metro_Area, data=train)

summary(reg_log)
```

When we log transform the regression, we see that the model is significant with a p-value of $< 2.2e-16$. However, after log-transforming the response the `sqft_lot` and `renovated1` variables became insignificant. We decided to run a General Linear F test to see if the model would be improved after removing these predictors. The null and alternate hypotheses are as follows:

$$
H_0: \beta_{\text{sqft\_lot}} = \beta_{\text{renovated}} = 0
$$

$$
H_a: \text{At least one of the coefficients in } H_0 \text{ is not equal to } 0
$$

```{r, echo=FALSE}
reg_log2 <- lm(log_price ~bedrooms + bathrooms + sqft_living + floors + waterfront + view + condition + grade + yr_built + Warm_Cold + Metro_Area, data=train)

anova(reg_log, reg_log2)
```

The F statistic from this test is 1.495, with a p-value of 0.2243. We fail to reject the null hypothesis, so there is little evidence of supporting the full model. We go with the reduced model over the full model.

#### Final Model: Removing Sqft_lot and Renovated

```{r, echo=FALSE}
summary(reg_log2)
```

##### Final Multicollinearity Check

```{r, echo=FALSE}
print(round(faraway::vif(reg_log2),3))
```

##### Final Assumption Tests

```{r echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
par(mfrow = c(2, 2))
plot(reg_log2)
```

We developed a final multiple linear regression model predicting $\log{(price)}$ using eleven predictors (`bedrooms`, `bathrooms`, `sq_living`, `floors`, `waterfront`, `view`, `condition`, `grade`, `yr_built`, `Warm_Cold`, and `Metro_Area`). All included variables were statistically significant in contributing to the model's predictive ability on the test data. A particular predictor that appears to have a large impact on the prices of houses in King County is whether it has a waterfront or not. The estimated difference in the log price between houses with and without a waterfront is $3.647e-01$. Other variables that have a large impact on the price of houses is the grade as well as if the house is located in the Metropolitan area. The final regression equation is as follows: $$
\begin{aligned}
\log(\text{price}) & = 19.47 \\
& - 0.0217 \times \text{bedrooms}\\
& + 0.08323 \times \text{bathrooms}\\
& + 0.0001827 \times \text{sqft\_living}\\
& + 0.06126 \times \text{floors}\\
& +0.3647 \times \text{waterfront}\\
& +0.04317 \times \text{view}\\
& +0.05147 \times \text{condition}\\
& +0.2225 \times \text{grade}\\
& - 0.004543 \times \text{yr\_built}\\
& + 0.03449 \times \text{Warm\_Cold}\\
& + 0.1151 \times \text{Metro\_Area}
\end{aligned}
$$

Since price is log-transformed, the interpretation of coefficients in context is like so: - For a 1-unit increase in bathrooms, there is an $e^{0.08323} = 1.086792$ increase, or $8.3$% - For a home with a waterfront, there is an $e^{0.3647} = 1.440082$ increase, or 44%, when compared to a home with no waterfront.

In this final model, multicollinearity is not a concern, with all VIF values below 5. Assumption tests showed improved residual behavior after the log transformation. The vertical spread of residuals appeared more constant across fitted values which passes the constant variance assumption (Assumption 2). Additionally, the red smoothed line overlaid on the residual plot is relatively flat and centered around zero, indicating that the linearity assumption (Assumption 1) is met. The final model achieved an adjusted R-squared of 0.6581, suggesting that approximately 66% of the variation in housing prices can be explained by the included predictors. Finally, with an F-statistic of 1890 and a p-value of less than $2.2e-16$, we conclude that the final model is highly statistically significant for predicting house prices in King County, Washington within the test data.

#### Checking for high leverage observations

```{r, echo=FALSE}
hii<-lm.influence(reg_log2)$hat 
ext.student<-rstudent(reg_log2) 
n<-nrow(Data)
p<-12
head(sort(hii[hii>2*p/n], decreasing = TRUE))
```

##### Externally Studentized Residuals

```{r, echo=FALSE}
ext.student[abs(ext.student)>3]
```

#### Detecting influential observations

##### DFFITS

```{r, echo=FALSE}
DFFITS<-dffits(reg_log2)
head(sort(DFFITS[abs(DFFITS)>2*sqrt(p/n)], decreasing = TRUE))
```

##### Cook's Distance

```{r, echo=FALSE}
COOKS<-cooks.distance(reg_log2)
COOKS[COOKS>1]
```

Although several observations exhibited high leverage and externally studentized residuals, we further assessed their overall influence on the model using DFFITS and Cook’s Distance. While a few points exceeded the DFFITS threshold, no observations had a Cook’s Distance greater than 1. Based on this, we concluded that no observations had a large enough impact on the regression results to be removed. Therefore, we decided to keep all observations in the final model.

# Section 6: Visualization of Good Quality

In this section, we analyze the available predictors of home quality using visualizations. Good quality is determined by a condition value greater than 3 and a grade value greater than 7. To better understand what quality consists of, we first created univariate visualizations showing the count of grades and conditions in our data set. As shown below, the highest counts of conditions and grades were low quality, but overall evenly spread.

```{r message=FALSE, warning=FALSE, echo = FALSE}
library(tidyverse)
library(datasets)
library(MASS)
library(ggplot2)
library(leaps) 
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
Data <-read.csv("kc_house_data.csv", header=TRUE)
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
## Reformat date
Data$date <- Data$date %>%
  substr(1,8) %>%
  as.Date(format = "%Y%m%d")

# Create variable for seasons
Data <- Data %>%
  mutate(
    month = as.numeric(format(date, "%m")),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    )
  )

Data <- Data %>%
  mutate(
    Warm_Cold = case_when(
      season == "Spring" | season == "Summer" ~ 1,
      season == "Fall"   | season == "Winter" ~ 0
    ),
    Warm_Cold = factor(Warm_Cold, levels = c(0,1))
  )

## Change record with 33 bedrooms to 3 (mistype)
Data[Data$id == 2402100895, "bedrooms"] <- 3

## Create binary variable for whether renovated
Data <- Data %>%
  mutate(renovated = if_else(yr_renovated == 0, 0, 1))

## Create binary variable for whether basement
Data <- Data %>%
  mutate(basement = if_else(sqft_basement == 0, 0, 1))

## Create variable for most recent year built
Data$recent.build <- pmax(Data$yr_built, Data$yr_renovated)

## Identify zipcodes within city limits
Data$zip_sub <- substr(Data$zipcode,2, 3) 
Data$Metro_Area <- ifelse(Data$zip_sub == '81', 1, 0)

## Create variable for time series
Data$days <- as.integer(Data$date - min(Data$date))

# Binary if good quality
Data$good_quality <- ifelse(Data$condition > 3 & Data$grade > 7, 1, 0)

#total rooms to reduce multicollinearity
Data$total_rooms <- Data$bedrooms + Data$bathrooms

## Change waterfront/zipcode to factors
Data$season <- factor(Data$season)
Data$waterfront <- factor(Data$waterfront, levels = c(0,1))
Data$zipcode <- factor(Data$zipcode)
Data$renovated <- factor(Data$renovated)
Data$Metro_Area <- factor(Data$Metro_Area)
Data$good_quality <- factor(Data$good_quality)
Data$basement <- factor(Data$basement)

## Filtering to make sure a house has a bedroom, bathroom, and lot/living area greater than 0
Data <- subset(Data, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample.data, ]
test<-Data[-sample.data, ]
```

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
library(scales) 
library(ggthemes)

barplotgrade <- ggplot(train, aes(x = factor(grade), fill = factor(grade > 7, levels = c(FALSE, TRUE), labels = c("Bad Grade", "Good Grade")))) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("Bad Grade" = "darkorange", "Good Grade" = "navyblue"), name = "Grade Category") +
  theme_fivethirtyeight() +
  labs(x = "Grade", y = "Count", title = "Count of Grades") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

barplotcondition <- ggplot(train, aes(x = factor(condition), fill = factor(condition > 3, levels = c(FALSE, TRUE), labels = c("Bad Condition", "Good Condition")))) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("Bad Condition" = "darkorange", "Good Condition" = "navyblue"), name = "Condition Category") +
  theme_fivethirtyeight() +
  labs(x = "Condition", y = "Count", title = "Count of Conditions") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

grid.arrange(barplotgrade, barplotcondition, ncol = 2)
```

Next, we examined how our binary variables, `Metro_Area` and `basement` associated with quality. Our team expected that homes in the city would have lower grades and conditions. As we can see below, high grades tend to be more common outside of the Seattle city limits. Additionally, high grades and conditions tend do be more common in homes with basements.

```{r echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
library(scales) 
library(ggthemes)

barplotgrade2 <- ggplot(train, aes(x = factor(grade), fill = Metro_Area)) +
  geom_bar(position = 'fill', alpha = 0.8) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), name = "Metro Area") +
  theme_fivethirtyeight() +
  labs(x = "Grade", y = "Proportion", title = "Proportion of Grades by Metro Area") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

barplotcondition2 <- ggplot(train, aes(x = factor(condition), fill = Metro_Area)) +
  geom_bar(position = 'fill', alpha = 0.8) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), name = "Metro Area") +
  theme_fivethirtyeight() +
  labs(x = "Condition", y = "Proportion", title = "Proportion of Conditions by Metro Area") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

barplotgrade3 <- ggplot(train, aes(x = factor(grade), fill = basement)) +
  geom_bar(position = 'fill', alpha = 0.8) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), name = "Basement") +
  theme_fivethirtyeight() +
  labs(x = "Grade", y = "Proportion", title = "Proportion of Grades by Basement") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

barplotcondition3 <- ggplot(train, aes(x = factor(condition), fill = basement)) +
  geom_bar(position = 'fill', alpha = 0.8) +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), name = "Basement") +
  theme_fivethirtyeight() +
  labs(x = "Condition", y = "Proportion", title = "Proportion of Conditions by Basement") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    legend.key.size = unit(0.5, "cm"))

grid.arrange(barplotgrade2, barplotcondition2, barplotgrade3, barplotcondition3, ncol = 2, nrow = 3)
```

Next, we shifted focus toward some of our quantitative predictors, including `yr_built`, `sqft_living`, and `price`. Our team hypothesized that homes that are more recently built would be of higher quality, and that larger homes would be associated with high quality. Based off the following plots, this appears to be true for square footage, but yr_built seems to have an opposite association, with a majority of good quality homes being built around 1975. Unsurprisingly, it seems that high prices are also associated with good quality homes.

```{r echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
densitytheme <- theme(plot.title = element_text(hjust = 0.5, size = 10),
  axis.title = element_text(size = 8),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9),
  legend.key.size = unit(0.2, "cm"))

densyr_built <- ggplot(train, aes(x = yr_built, fill = good_quality))+
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  labs(x = "Year Built", y = "Density", title = "Year Built by Good Quality")+
  densitytheme

denssqft_living <- ggplot(train, aes(x = sqft_living, fill = good_quality))+
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  labs(x = "Square Feet of Living Space", y = "Density", title = "Sqft_Living by Good Quality")+
  densitytheme

densprice <- ggplot(train, aes(x = price, fill = good_quality))+
  geom_density(alpha = 0.8)+
  theme_fivethirtyeight()+
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"),labels = c("No", "Yes"), name = "Good Quality?") +
  scale_x_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  labs(y = "Density", title = "Price by Good Quality")+
  densitytheme

grid.arrange(densyr_built, denssqft_living, densprice)
```

Lastly, we again visualized the relationship between `sqft_living` and `price`, separated by `good_quality`, with the orange points being the points of focus. It is evident from the graphs that price and sqft_living have a positive relationship. While there is much overlap between the two groups, our visualizations indicate that price and sqft_living are both higher for homes with good quality.

```{r message=FALSE, warning=FALSE, echo = FALSE}
ggplot(train, aes(x = sqft_living, y = price, color = good_quality))+
  geom_point(data = transform(train, good_quality = NULL), color = 'navyblue')+
  geom_point(color = 'darkorange') +
  theme_fivethirtyeight() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
  facet_wrap(vars(good_quality))+
  labs(title = "Sqft_Living against Price, by Good Quality", x = "Sqft_Living")+
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M"), name = "Price (in Millions)", limits = c(0, 4000000)) +
  theme(plot.title = element_text(hjust = 0.5, size = 10),
  axis.title = element_text(size = 8),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9),
  legend.key.size = unit(0.2, "cm"))

```

# Section 7: Logistic Regression

```{r, echo = FALSE, results=FALSE}
library(tidyverse)

#Read data and split it into test/train
Data<-read.csv("kc_house_data.csv", sep=",", header=TRUE)
set.seed(6021)
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F) #split data on a 50/50
kc.df<-Data[sample.data, ] #training data
kc.dftest<-Data[-sample.data, ] #test data

#Use train for all regression, test data for predictability calcs
#remove duplicated rows
kc.df <- kc.df[!duplicated(kc.df), ]

## reformat date
kc.df$date <- kc.df$date %>%
  substr(1,8) %>%
  as.Date(format = "%Y%m%d")

#add season based on month in date
kc.df <- kc.df %>%
  mutate(
    month = as.numeric(format(date, "%m")),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    )
  )

# houses should not have 0 bedrooms, bathrooms, or sqft
kc.df <- subset(kc.df, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)

#factor waterfront 
kc.df$is_waterfront <- factor(kc.df$waterfront)

## factors for waterfront and zipcode
kc.df$waterfront <- factor(kc.df$waterfront, levels = c(0,1))
kc.df$zipcode <- factor(kc.df$zipcode)

## change 33 bedrooms record to 3
kc.df[kc.df$id == 2402100895, "bedrooms"] <- 3
kc.df[kc.df$id == 2402100895,]

## renovated/not renovated column
kc.df <- kc.df %>%
  mutate(renovated = if_else(yr_renovated == 0, 0, 1))

## most recent year of build
kc.df$recent.build <- pmax(kc.df$yr_built, kc.df$yr_renovated)

## Define metropolitan ZIP codes
kc.df$zip_sub <- substr(kc.df$zipcode,2, 3) 
kc.df$Metro_Area <- ifelse(kc.df$zip_sub == '81', 1, 0)

## filtering to make sure a house has a bedroom, bathroom, and lot/living area greater than 0
subset(kc.df, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)

#total rooms to reduce multicollinearity
kc.df$total_rooms <- kc.df$bedrooms + kc.df$bathrooms

#per section 7 requirements 
kc.df$good_quality <- ifelse(kc.df$condition > 3 & kc.df$grade > 7, 1, 0)

#presence of basement
kc.df$basement <- ifelse(kc.df$sqft_basement > 0, 1,0)

summary(kc.df) 

```

```{r, echo = FALSE, results = FALSE}
# Add same columns to the test dataset
kc.dftest <- kc.dftest[!duplicated(kc.dftest), ]

## reformat date
kc.dftest$date <- kc.dftest$date %>%
  substr(1,8) %>%
  as.Date(format = "%Y%m%d")

#add season based on month in date
kc.dftest <- kc.dftest %>%
  mutate(
    month = as.numeric(format(date, "%m")),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    )
  )

# houses should not have 0 bedrooms, bathrooms, or sqft
kc.dftest <- subset(kc.dftest, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)

#factor waterfront 
kc.dftest$is_waterfront <- factor(kc.dftest$waterfront)

## factors for waterfront and zipcode
kc.dftest$waterfront <- factor(kc.dftest$waterfront, levels = c(0,1))
kc.dftest$zipcode <- factor(kc.dftest$zipcode)

## change 33 bedrooms record to 3
kc.dftest[kc.dftest$id == 2402100895, "bedrooms"] <- 3
kc.dftest[kc.dftest$id == 2402100895,]

## renovated/not renovated column
kc.dftest <- kc.dftest %>%
  mutate(renovated = if_else(yr_renovated == 0, 0, 1))

## most recent year of build
kc.dftest$recent.build <- pmax(kc.dftest$yr_built, kc.dftest$yr_renovated)

## Define metropolitan ZIP codes
kc.dftest$zip_sub <- substr(kc.dftest$zipcode,2, 3) 
kc.dftest$Metro_Area <- ifelse(kc.dftest$zip_sub == '81', 1, 0)

## filtering to make sure a house has a bedroom, bathroom, and lot/living area greater than 0
subset(kc.dftest, bedrooms > 0 & bathrooms > 0 & sqft_living > 0 & sqft_lot > 0)

#total rooms to reduce multicollinearity
kc.dftest$total_rooms <- kc.dftest$bedrooms + kc.dftest$bathrooms

#per section 7 requirements 
kc.dftest$good_quality <- ifelse(kc.dftest$condition > 3 & kc.dftest$grade > 7, 1, 0)

#presence of basement
kc.dftest$basement <- ifelse(kc.dftest$sqft_basement > 0, 1,0)
```

Our team used a logistic regression model to model whether a home in King County, Washington is of good quality, where good quality was defined as homes having a condition score greater than 3 and a grade greater than 7. Our initial model included a broad set of predictors based on factors that influence house quality which included total rooms, price of the home, sqft_living, sqft_lot, floors, season, waterfront, view, condition, grade, yr_built, renovated, Metro_Area, and basement. Our aim was to start with many factors and whittle the model down to those that were useful. The following figure is the R output of the first model.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#the full logisitic model using training data
kc.dfmodel <- glm(good_quality~ total_rooms + price+sqft_living + sqft_lot+floors +season+ waterfront +view + condition + grade +yr_built + renovated + Metro_Area+ basement, data=kc.df, family = "binomial")
summary(kc.dfmodel)
```

After fitting the full model, we improved it through several steps. Based on the logistic regression summary, we found that variables sqft_lot, waterfront, floors, renovated, and seasons had high p-values that were greater than .05, and as such, were not statistically significant. Next, we decided to fit a reduced model with all of the variables that were statistically significant with a p-value less than 0.001. The resulting model had the variables total_rooms, price, sqft_living, view, condition, grade, yr_built, Metro_Area, and basement. The R output for this reduced model is as follows:

```{r, echo = FALSE}
# reduced model with training data summary
reduced_kc.df<-glm(good_quality~ total_rooms + sqft_living + view + price+condition + grade +yr_built + Metro_Area+ basement, data=kc.df, family = "binomial")
summary(reduced_kc.df)
```

Based on the R output, the logistic regression equation is the following: $$
\begin{aligned}
\log\left(\frac{\pi}{1 - \pi}\right) &= 5.472 \\
& + 0.1856 \times \text{total\_rooms} \\
& - 0.0005659 \times \text{sqft\_living} \\
&+ 0.2058 \times \text{view} \\
& - 0.0000007687 \times \text{price} \\
& + 3.231 \times \text{condition} \\
& + 1.931 \times \text{grade} \\
& - 0.01745 \times \text{yr\_built} \\
& - 1.267 \times \text{Metro\_Area} \\
& + 0.7064 \times \text{basement}
\end{aligned}
$$

Our team noted that all variables are statistically significant. To be sure that the reduced model is truly more useful, we conducted a likelihood ratio test comparing the full and reduced models with the null hypothesis supporting the reduced model and the alternative hypothesis supporting the full model. The null and alternate hypotheses are as follows:

$$
\begin{aligned}
H_0: &\ \beta_{\text{sqft\_lot}} = \beta_{\text{waterfront}} = \beta_{\text{floors}} = \beta_{\text{renovated}} = \beta_{\text{seasons}} = 0 \\
H_a: &\ \text{At least one of the coefficients in } H_0 \text{ is not 0}
\end{aligned}
$$

```{r, echo= FALSE, results = FALSE}
TS<-reduced_kc.df$deviance-kc.dfmodel$deviance
1-pchisq(TS,7)
```

From the likelihood ratio test, we obtained a p-value of 0.6510558. As this is greater than 0.05, we failed to reject the null hypothesis and dropped the predictors, as there was not enough evidence that they were useful to the model.

Using this reduced model we concluded many interesting findings. First, we found that living space (`sqft_living`) and total rooms were positively associated with larger living spaces. Additionally, homes with more rooms were more likely to be of good quality. We also found that based on `yr_built,` newer homes were slightly more likely to be of good quality. We also found that higher view scores were more likely to be classified as good quality. These findings can be summarized by the visualizations in the following figure:

```{r echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
densitytheme <- theme(
  plot.title = element_text(hjust = 0.5, size = 10),
  axis.title = element_text(size = 8),
  legend.text = element_text(size = 8),
  legend.title = element_text(size = 9),
  legend.key.size = unit(0.2, "cm"))

vis1 <- ggplot(kc.df, aes(x = total_rooms, fill = factor(good_quality))) +
  geom_density(alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  labs(title = "Total Rooms Distribution by Good Quality", x = "Total Rooms", y = "Density") +
  densitytheme

vis2 <- ggplot(kc.df, aes(x = factor(Metro_Area), fill = factor(good_quality))) +
  geom_bar(position = "fill", alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  labs(title = "Good Quality Homes by Metro Area", x = "Metro Area (0 = No, 1 = Yes)", y = "Proportion") +
  densitytheme

vis3 <- ggplot(kc.df, aes(x = yr_built, fill = factor(good_quality))) +
  geom_density(alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  labs(title = "Year Built Distribution by Good Quality", x = "Year Built", y = "Density") +
  densitytheme

vis4 <- ggplot(kc.df, aes(x = sqft_living, fill = factor(good_quality))) +
  geom_density(alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_color_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  labs(title = "Square Feet of Living Space by Good Quality", x = "Square Feet", y = "Density") +
  densitytheme

vis6 <- ggplot(kc.df, aes(x = factor(view), fill = factor(good_quality))) +
  geom_bar(position = "fill", alpha = 0.8) +
  theme_fivethirtyeight() +
  scale_fill_manual(values = c("0" = "darkorange", "1" = "navyblue"), labels = c("No", "Yes"), name = "Good Quality?") +
  labs(title = "Good Quality Homes by View Type", x = "View", y = "Proportion") +
  densitytheme

gridExtra::grid.arrange(vis1, vis6, vis3, vis4, ncol = 2, nrow = 2)
```

After fitting an appropriate logistic regression model, our team also wanted to measure it's predictive performance. We by finding the area under the ROC curve, using a test data set. Based on the ROC curve, the curve is far from the diagonal and closer to the top left of the plot, which indicates that the model is classifying observations accurately. We also found the area under the curve to be 0.95. Since this value is greater than 0.5, we can conclude that our model performs better than random guessing. The following plot shows our the ROC curve for our model:

```{r, echo = FALSE}
#ROC and AUC using test data
library(ROCR)

# Get predicted probabilities
prob <- predict(reduced_kc.df, newdata = kc.dftest, type = "response")

pred <- ROCR::prediction(prob, kc.dftest$good_quality)
roc.res <- ROCR::performance(pred, measure="tpr", x.measure="fpr")
plot(roc.res, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red")
```

```{r, echo = FALSE, results=FALSE}
auc<-ROCR::performance(pred, measure = "auc")
auc@y.values
```

Our team took additional measures to evaluate model performance, including creating a confusion matrix using a 0.5 probability threshold and a 0.3 probability threshold. Our confusion matrix with a 0.5 probability threshold achieved 89.5% accuracy, an error rate of 10.5%, a true negative rate of 95.5%, and a true positive rate of 45.3%. However, we observed that lowering the threshold to 0.3 significantly improved the model's ability to correctly classify good quality homes, increasing the true positive rate to 66.5% with only a small increase in the true negative rate to 92.9% while decreasing the error rate to 10.3%.

**Confusion Matrix using 0.5 Probability Threshold:**

```{r, echo = FALSE}
#Confusion matrix using 0.5 threshold
table(kc.dftest$good_quality, prob>0.5)
#(591+9066)/(9066+429+709+591) #accuracy
#TPR<-591/(713+591)
#TPR
#TNR<-(9066/(9066+426))
#TNR
```

**Confusion Matrix using 0.3 Probability Threshold:**

```{r, echo = FALSE}
#Confusion matrix using 0.3
table(kc.dftest$good_quality, prob>0.3)
#(8812+851)/(8812+851+680+453) #accuracy
#TPR<- 867/(437+867)
#TPR
#TNR<- 8818/(8818+674)
#TNR
```

Our logistic regression model effectively predicts whether a home in King County, Washington is of good quality based on a model with strong, statistically significant predictors generalizing living space, location, and quality of the home. The model generalizes well to new data with a low error rate and high accuracy of identifying good quality homes in the area. Potential home buyers can use the results of our model to gain comfort that their new homes will be of good quality.
